### 2024-03-27 Manuscript.Rmd

---
title: "Manuscript Figures"
author: "Charlie Bayne"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        toc: true
        toc_depth: 6
        toc_float:
            collapsed: true
            smooth_scroll: true
        code_folding: hide
        highlight: tango
        df_print: paged
---

```{r setup, include=FALSE, warning= FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)

knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,cache = TRUE)

#Setting ggplot2 theme, this isn't working for some reason
theme_set(
  ggprism::theme_prism()+
  theme(plot.title = element_text(hjust = 0.5))
  )

#Setting ggplot2 theme, this one works
#theme_set(
 # theme_classic()+
  #theme(plot.title = element_text(hjust = 0.5))
  #)
```

```{r}
#Loading in mapping data that will allow us to go from Uniprot ID to gene ID. 
# Will use this to fix the labels. 
# Next time make sure to have fragpipe give gene level data instead of by uniprot id.
protein_to_gene = readr::read_csv("00_r_input_data/protein_id_to_gene.csv")
```

# Figure 1 - Overview of proteomics and metabolomics data

## A - Description of the overall study.

This figure was generated in Biorender.

## B- Characteristics of patient samples.

```{r,fig.height = 2.5, fig.width = 8.268}
### Figure 1b ###

# Reading in the cleaned clinical metadata to start to make plots
clinical_metadata = readr::read_csv("00_r_input_data/clinical_metadata.csv")

p0 = clinical_metadata %>% 
  ggplot(aes(condition,fill = condition))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d()+
  ggtitle("Condition")+
  ggprism::theme_prism(base_size = 6)

p0

# Defining the columns that we want to show patient charachteristics for
columns_to_look_at = c("gender", "charleson_index", "bacteremia_duration","gender","day_of_blood_draw","sensitive_to_vancomycin","condition","death_during_admission")

# Filtering the data to the relevant columns
sel = clinical_metadata %>% 
  select(sample_id,columns_to_look_at,pathogen) %>% 
  filter(!is.na(pathogen))

# Plotting Gender by Pathogen
p1 = sel %>% 
  ggplot(aes(gender,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Gender") +
  ggprism::theme_prism(base_size = 6)

#Plotting Charleson index by Pathogen
p2 = sel %>% 
  ggplot(aes(charleson_index,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Charleson Index")+
  ggprism::theme_prism(base_size = 6)

# Bacteremia Durration

p3 = sel %>% 
  ggplot(aes(bacteremia_duration,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Bacteremia Duration")+
  ggprism::theme_prism(base_size = 6)

# Day of Blood Draw

p4 = sel %>% 
  ggplot(aes(day_of_blood_draw,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Day of Blood Draw")+
  ggprism::theme_prism(base_size = 6)

# Sensitivity to Vancomycin

p5 = sel %>% 
  ggplot(aes(sensitive_to_vancomycin,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Vancomycin Sensitivity")+
  ggprism::theme_prism(base_size = 6)

# Sensitivity to Vancomycin

p6 = sel %>% 
  ggplot(aes(death_during_admission,fill = pathogen))+
  geom_histogram(stat = "count")+
  scale_fill_viridis_d(end = 0.5)+
  ggtitle("Death During Admission")+
  ggprism::theme_prism(base_size = 6)

f1b = ggpubr::ggarrange(p1,p2,p3,p4,p5,p6,common.legend = TRUE,legend = "none",ncol = 6,font.label = list(size = 14))

f1b

ggsave("02_figures/f1_overview_of_study/f1b.pdf",plot = f1b, height = 2.5, width = 8.268,units = "in")
```

## C- Overall clustering of proteomoics data.

```{r,fig.height = 8, fig.width = 16}
### Figure 1C ###

#Loading in mapping file
mapping = readxl::read_excel("../proteomics/sample_mapping.xlsx")

# Loading in the clinical metadata 
cmd = read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  select(-condition)

mapping = inner_join(mapping,cmd,by = c("sample_name" = "sample_id"))

#Loading in data
data = read_delim("../proteomics/fragpipe_output_fixed_labels/tmt-report/abundance_gene_MD.tsv")

#Converting to long format
long = data %>% 
  pivot_longer(6:length(.),names_to = "sample_name")

#Combining data and mapping
all = inner_join(mapping,long,by="sample_name")

#Converting into a matrix
mat = all %>% 
  dplyr::select(-ProteinID,-NumberPSM,-ReferenceIntensity,-MaxPepProb) %>% 
  pivot_wider(values_from = value,names_from = Index)

#Extracting only the data
dat = mat %>% 
  dplyr::select(81:length(.)) %>% 
  dplyr::select_if(~ !any(is.na(.)))

#Extracting only the metadata
md = mat %>% 
  dplyr::select(1:81)

#setting row names so we can look at the labels if need be
rownames(dat) <- paste0(md$sample_name)

# Calculating distances
d = dist(dat)

# Showing as a dendrogram
dend = as.dendrogram(hclust(d, method = "ward.D2"))

library(circlize)
library(dendextend)

# adding color by the type of infection
colors_to_use = viridis::viridis(3)[as.numeric(as.factor(md$condition))]

# getting everything in the correct order 
colors_to_use = colors_to_use[order.dendrogram(dend)]

# adding mortality
colors_to_use_2 = c("#000000","#FF0000")[as.numeric(as.factor(md$death_during_admission))]

colors_to_use_2 = colors_to_use_2[order.dendrogram(dend)]

# adding color to the dendrogram branches
dend = branches_color(dend,col = colors_to_use) %>% 
  set("branches_lwd", 8)

labels_colors(dend) = colors_to_use_2

#this is in circular form

pdf("02_figures/f1_overview_of_study/f1c.pdf")
circlize_dendrogram(dend,
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE)
dev.off()

circlize_dendrogram(dend,
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE)
```

## D- Overall clustering of meatabolomics data.

```{r,fig.height = 8, fig.width = 16}
### Figure 1 D ###

## Reading in data ##
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

##Why do we have different numbers of healthy?

test = normalized_data %>% 
  select(sample_id,condition) %>% 
  distinct() %>% 
  group_by(condition) %>% 
  summarise(n = n())

## Transforming the data ##
norm_wide = normalized_data %>%
  select(-row_m_z,-row_retention_time,-value) %>% 
  tidyr::pivot_wider(names_from = row_id,values_from = norm_value)

norm_dat = norm_wide %>% 
  dplyr::select(78:length(.))

norm_md = norm_wide %>% 
  dplyr::select(1:77)

rownames(norm_md) = norm_md$sample_id

## Plotting the overal clustering ##

rownames(norm_dat) = norm_md$sample_id

d = dist(norm_dat)

dend = as.dendrogram(hclust(d, method = "ward.D2"))

#dend = as.dendrogram(hclust(d))

library(circlize)
library(dendextend)

# let's add some color:
colors_to_use = viridis::viridis(3)[as.numeric(as.factor(norm_md$condition))]

#getting in the same order as dendrogram
colors_to_use = colors_to_use[order.dendrogram(dend)]

# adding mortality
colors_to_use_2 = c("#000000","#FF0000")[as.numeric(as.factor(norm_md$death_during_admission))]

colors_to_use_2 = colors_to_use_2[order.dendrogram(dend)]

# Now we can color the brancehes
dend = branches_color(dend,col = colors_to_use,) %>% 
  set("branches_lwd", 8)

labels_colors(dend) = colors_to_use_2

pdf("02_figures/f1_overview_of_study/f1d.pdf")
circlize_dendrogram(dend,
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE)
dev.off()

circlize_dendrogram(dend,
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE)
```

# Figure 2 - Host Response Bacteremias Proteomics

## A- Volcano plot of healthy vs Infected EcB

### Data Processing
```{r}
# Read MSstats.csv file.
data = read_csv("00_r_input_data/msstats.csv")
#Note that the condition of the bridge channel needs to be labeled as Norm
annotation = read_csv("00_r_input_data/MSstatsTMT_annotation.csv")

mstats = MSstatsTMT::PhilosophertoMSstatsTMTFormat(input = data,annotation)

stats = MSstatsTMT::PhilosophertoMSstatsTMTFormat(input = data,annotation)

# use MSstats for protein summarization
quant.msstats = MSstatsTMT::proteinSummarization(stats,
                                      method="msstats",
                                      global_norm=TRUE,
                                      reference_norm=TRUE,
                                      remove_norm_channel = TRUE,
                                      remove_empty_channel = TRUE)

# Taking the pl_data and formatting appropriately
pl_data = quant.msstats$ProteinLevelData %>% 
  tidyr::separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>%
  # Converting to gene name
  inner_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  mutate(sample_id = paste0(Mixture,".",Channel))

write_csv(pl_data,"01_r_processed_data/quant_msstats_protein_level_data.csv")

write_csv(quant.msstats$FeatureLevelData,"01_r_processed_data/quant_msstats_feature_level_data.csv")

### Preforming the statistics ### 
test.pairwise = MSstatsTMT::groupComparisonTMT(quant.msstats, moderated = TRUE)

## All groups ## 
res = test.pairwise$ComparisonResult %>% 
  separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|")

write_csv(res,"01_r_processed_data/groupComparison_results.csv")

## Faecalis v Healthy
faecalis_v_healthy = res %>% 
  filter(Label == "Faecalis vs Healthy")

write_csv(faecalis_v_healthy,"01_r_processed_data/faecalis_v_healthy.csv")

## Faecium v Healthy 
healthy_v_faecium = res %>% 
  filter(Label == "Healthy vs Faecium")

write_csv(healthy_v_faecium,"01_r_processed_data/healthy_v_faecium.csv")

## Faecalis v Faecium ##
faecalis_v_faecium = res %>% 
  filter(Label == "Faecalis vs Faecium")

write_csv(faecalis_v_faecium,"01_r_processed_data/faecalis_v_faecium.csv")
```

### How many are differentially abundant?

```{r}
infected_vs_healthy_plot %>% 
  mutate(sign = case_when(log2FC < 0 ~ "negative",
                          TRUE ~ "positive")) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  group_by(sign) %>% 
  summarise(n = n())

```

```{r}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

both_sig = infected_vs_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(uniprotid)

# Defining significant proteins in faecalis
faecalis_sig = faecalis_v_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(uniprotid)

# Defining significant proteins in faecium
faecium_sig = healthy_v_faecium %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(uniprotid)

library(VennDiagram)

venn.diagram(
        x = list(both_sig, faecalis_sig, faecium_sig),
        category.names = c("E.facalis + E.faecium" , "E.faecalis " , "E.faecium"),
        filename = "../Manuscript/02_figures/f2_infection_v_healthy_proteomics/f2a-venn_diagram.svg",
        output=TRUE,

        # Output features
        imagetype="svg" ,
        height = 5 , 
        width = 5 , 
        units = "in",

        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,

        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",

        # Set names
        cat.cex = 0.6,
        cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135),
        cat.dist = c(0.055, 0.055, 0.085),
        cat.fontfamily = "sans",
        rotation = 1
)

```

## B - Heat map by GO Terms

### Processing/ Generation of related supplementary figures

##### EcB

###### Up.

Network analysis of proteins that are shared between faecalis and faecium relative to healthy when considered alone and increased in infected samples

```{r,fig.height = 11, fig.width = 8}
##Loading all the go stuff

#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Extracting the uniprotids up in infected and shared

  infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid"= "ProteinID"))

shared_up_in_infected = infected_vs_healthy %>% 
  filter(log2FC > 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_up_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

up_ECB = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                circular = FALSE,
                                categorySize="pvalue",
                                cex_label_category = 0.9,
                                showCategory = 100,
                                cex_label_gene = 0.5
                                )+
  theme(legend.position = "none")

up_ECB

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_ECB.pdf",up_ECB,height=11,width=8)
```

###### Down

```{r,fig.height = 11, fig.width = 8}
##Loading all the go stuff

#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
    # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Extracting the uniprotids up in infected and shared
shared_down_in_infected = infected_vs_healthy %>% 
  filter(log2FC < 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_down_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

down_EcB = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

down_EcB

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_EcB.pdf",down_EcB,height =11, width = 18)
```
##### Faecalis

###### Up

```{r,fig.height = 11, fig.width = 8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid"= "ProteinID"))

# Extracting the uniprotids up in infected and shared
up_in_faecalis = faecalis_v_healthy %>% 
  filter(log2FC >= 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

up_faecalis = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                circular = FALSE,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

up_faecalis

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecalis.pdf",up_faecalis,height=11,width=8)
```

###### Down

```{r,fig.height = 11, fig.width=8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Extracting the uniprotids up in infected and shared
down_in_faecalis = faecalis_v_healthy %>% 
  filter(log2FC <= 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

down_faecalis = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                circular = FALSE,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

down_faecalis

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecalis.pdf",down_faecalis,height=11,width=8)
```

##### Faecium

###### Up

```{r,fig.height = 11, fig.width = 8}

healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid"= "ProteinID"))

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Extracting the uniprotids up in infected and shared
# Extracting the uniprotids up in infected and shared
up_in_faecium = healthy_v_faecium %>% 
  filter(log2FC <= 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

up_faecium = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                circular = FALSE,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

up_faecium

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecium.pdf",up_faecium,height=11,width=8)
```

###### Down
```{r,fig.height = 11, fig.width = 8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Extracting the uniprotids up in infected and shared
down_in_faecium = infected_vs_healthy_plot %>% 
  filter(log2FC <= 0) %>% 
  filter(color == "sig_in_faecium") %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

down_faecium = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                circular = FALSE,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

down_faecium

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecium.pdf",down_faecium,height=11,width=8)

```

##### Staph Aureus

```{r,eval=FALSE}
#Loading in data, getting GO annotations
staph_data = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1)

ids = staph_data %>% 
  pull(Protein)

# Getting GO annotations by hitting Uniprots API
anot_go = GLabR::annotate_uniprot_single(ids,columns = "accession,id,gene_primary,go,go_c,go_p,go_f,go_id")

write_csv(anot_go,"01_r_processed_data/staph_go_annotations.csv")

```

###### Up

```{r,fig.height = 11, fig.width= 8}
#Loading in data,
staph_data = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1)
anot_go = read_csv("01_r_processed_data/staph_go_annotations.csv") %>% 
  select(Entry = Gene.Names..primary.,Gene.Ontology.IDs) %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs))
    # Going from Protein ID to gene ID
 # left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>%

sig_up = staph_data %>% 
  filter(Infection_pvalue <= 0.05) %>% 
  filter(Infection_FC > 1) %>% 
  pull(Gene)

sig_down = staph_data %>% 
  filter(Infection_pvalue <= 0.05) %>% 
  filter(Infection_FC < 1) %>% 
  pull(Gene)

background = staph_data %>% 
  pull(Gene)

#usng enricher
TERM2GENE = anot_go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_up,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

up_staph = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")
up_staph

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_staph.pdf",up_staph,height=11,width=8)
```

###### Down

```{r,fig.height = 11, fig.width= 8}
# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_down,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

down_staph = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

down_staph

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_staph.pdf",down_staph,height=11,width=8)

```

#### Plotting Heatmap

```{r,fig.height =10, fig.width = 12}

## Enterococcal Bacteremia 
up_in_EcB_data = up_ECB$data[1:32,] %>% 
   mutate(type = "EcB")

down_in_EcB_data = down_EcB$data[1:16,] %>% 
  mutate(size = size * -1) %>% 
  mutate(type = "EcB")

## E. faecalis

up_in_faecalis_data = up_faecalis$data[1:24,] %>% 
  mutate(type = "E. faecalis")

down_in_faecalis_data = down_faecalis$data[1:24,] %>% 
  mutate(type = "E. faecalis") %>% 
  mutate(size = size * -1)

## E. faecium
up_in_faecium_data = up_faecium$data[1:9,] %>% 
  mutate(type = "E. faecium")

down_in_faecium_data = down_faecium$data[1:8,] %>% 
  mutate(type = "E. faecium") %>% 
  mutate(size = size * -1)

## Staph Aureus

up_staph_data = up_staph$data[1:8,] %>% 
  mutate(type = "S. aureus")

down_staph_data = down_staph$data[1:19,] %>% 
  mutate(size = size * -1) %>% 
  mutate(type = "S. aureus")

# Combining all of the data

all = bind_rows(up_in_EcB_data,
                down_in_EcB_data,
                up_in_faecalis_data,
                down_in_faecalis_data,
                up_in_faecium_data,
                down_in_faecium_data,
                up_staph_data,
                down_staph_data)

# plotting the heatmap 
p1 = all %>% 
  ggplot(aes(reorder(name,size),type,fill = size))+
  geom_tile()+
  coord_flip()+
  scale_fill_gradient2(
  name = waiver(),
  high = "red",
  low = "green",
  mid = "white",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) + 
  xlab("GO term")+
  ylab("Bacteremia Type")+
  ggprism::theme_prism(base_size = 10)

f2b

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2b.pdf",height = 10, width = 12)

```

## ? - Venn Diagram

### Up in infected
```{r,fig.height = 3.8, fig.width=4}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

# Defining significant proteins in faecalis
faecalis_sig = faecalis_v_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC<0) %>% 
  pull(uniprotid)

# Defining significant proteins in faecium
faecium_sig = healthy_v_faecium %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC > 0) %>% 
  pull(uniprotid)

# found to be sig in both faecalis and faecium
enterococcus_sig = infected_vs_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC < 0)%>% 
  pull(uniprotid)

# Loading staph_data
staph_sig = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1) %>% 
  filter(Infection_pvalue <= 0.05,
         Infection_FC < 1) %>% 
  pull(Protein)

# Generating venn diagram plot

library(RColorBrewer)
myCol <- brewer.pal(4, "Pastel2")

library(VennDiagram)

venn.diagram(
        x = list(enterococcus_sig, faecalis_sig, faecium_sig, staph_sig),
        category.names = c("EcB" , "E.faecalis " , "E.faecium","S. aureus"),
        filename = '02_figures/f2_infection_v_healthy_proteomics/up_in_infected_venn_diagram.png',
        output=TRUE,

        # Output features
        imagetype="png" ,
        height = 1100 , 
        width = 1100 , 
        resolution = 300,
        compression = "lzw",

        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,

        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",

)
```

### Up in Healthy

```{r,fig.height = 3.8, fig.width=4}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

# Defining significant proteins in faecalis
faecalis_sig = faecalis_v_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC>0) %>% 
  pull(uniprotid)

# Defining significant proteins in faecium
faecium_sig = healthy_v_faecium %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC < 0) %>% 
  pull(uniprotid)

# found to be sig in both faecalis and faecium
enterococcus_sig = infected_vs_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC > 0)%>% 
  pull(uniprotid)

# Loading staph_data
staph_sig = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1) %>% 
  filter(Infection_pvalue <= 0.05,
         Infection_FC > 1) %>% 
  pull(Protein)

# Generating venn diagram plot

library(RColorBrewer)
myCol <- brewer.pal(4, "Pastel2")

library(VennDiagram)

venn.diagram(
        x = list(enterococcus_sig, faecalis_sig, faecium_sig, staph_sig),
        category.names = c("EcB" , "E.faecalis " , "E.faecium","S. aureus"),
        filename = '02_figures/f2_infection_v_healthy_proteomics/up_in_healthy_venn_diagram.png',
        output=TRUE,

        # Output features
        imagetype="png" ,
        height = 1100 , 
        width = 1100 , 
        resolution = 300,
        compression = "lzw",

        # Circles
        lwd = 2,
        lty = 'blank',
        fill = myCol,

        # Numbers
        cex = .6,
        fontface = "bold",
        fontfamily = "sans",

)

```
## C- Evaluation of Proteomic Biomarkers

### Generating top 10 biomarkers

```{r,fig.height=8,fig.width = 8}
#reading in the data
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

## Getting into a format ot use with EFS ##
clean = pl %>% 
  dplyr::group_by(Mixture,Gene,Channel,Condition) %>% 
  dplyr::summarise(sum = sum(Abundance)) %>% 
  ungroup()

infected_healthy_l = clean %>% 
  mutate(classlabels = case_when(Condition %in% c("Faecalis","Faecium") ~ 1,
                                   TRUE ~ 0)) %>% 
  mutate(classlabels = as.logical(classlabels)) %>% 
  # - messes up EFS for some reason, very annoying. 
  mutate(Gene = gsub("-","_",Gene))

infected_healthy_w = infected_healthy_l %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = sum) %>% 
  mutate(sampleid = paste0(Mixture,".",Channel)) %>% 
  dplyr::select(-Mixture,-Channel,-Condition,-sampleid) %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame()

set.seed(2)

## Running EFS ##
ensemble_result = EFS::ensemble_fs(infected_healthy_w,1,cor_threshold = 0)

## Getting EFS Result in Plotable Format
long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "Gene") %>% 
  #converting genes back to proper notation. 
  mutate(Gene = gsub("_","-",Gene))

# Ranking Biomarkers
top = long %>% 
  group_by(Gene) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

# Extracting the top 10 ranked biomarkers 
topn = top[1:10,] %>% 
  pull(Gene)

# Filtering the EFS data for plotting purposes
filtered = long %>% 
  filter(Gene %in% topn)

# # Generating the EFS plot. 
# f2d = filtered %>% 
#   ggplot(aes(reorder(Gene,value),value,fill = model))+
#   geom_col()+
#   coord_flip()+
#   viridis::scale_fill_viridis(discrete = TRUE)+
#   ggtitle("Top 10 Performing Protein Biomarkers")+
#   xlab("Protein Biomarkers")
# 
# 
# f2d

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2d.pdf",f2d,height = 8, width = 8)

# Writting to file to use in supplementary figures. 
topn_proteomics = tibble(Gene = topn)

write_csv(topn_proteomics,"01_r_processed_data/top10_infected_v_healthy_proteomics_biomarkers.csv")

```

### Plotting

```{r,fig.height=8,fig.width=8}
# Reading in data
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

# Processing data 
ROC_data = dplyr::filter(protein_sum,Gene %in% topn) %>% 
  dplyr::mutate(Condition = case_when(Condition %in% c("Faecalis","Faecium") ~ "infected",
                                      TRUE ~ Condition))

source("functions.R")

bottom_plot = make_roc_proteomics(ROC_data,topn[1:2],ncol = 2)

top_plot = make_violin_plot_proteomics(ROC_data,topn[1:2], ncol = 2,pname = "p.signif")

f2c = ggpubr::ggarrange(top_plot,bottom_plot,nrow = 2)
ggsave("02_figures/f2_infection_v_healthy_proteomics/f2c.pdf",f2c,height = 8, width = 8)
```

# Figure 3- Metabolomic differences between Healthy and Infected EcB

## Processing Metabolomic data

```{r}
### Unnormalized data ###
#reading and formatting the metabolomics data
data = read_csv("00_r_input_data/mzmine_output/features_quant.csv") %>% 
  janitor::clean_names() %>% 
  select(1:3,14:length(.)) %>% 
  pivot_longer(4:length(.),names_to = "metabolomics_id") %>% 
  mutate(metabolomics_id = gsub("_mz.*","",metabolomics_id),
         metabolomics_id = toupper(metabolomics_id)) %>% 
  #need to add the RID prefix because having numeric column names (which we will have after pivoting the data) messes up lots of the functions. 
  mutate(row_id = paste0("RID_",row_id))

# Reading in the mapping 
map = read_csv("00_r_input_data/metabolomics_mapping.csv")

md = read_csv("00_r_input_data/clinical_metadata.csv")

# combining the mapping and metadata
map_md = inner_join(map,md,by= "sample_id")

# Combining everything
all = inner_join(data,map_md,by = "metabolomics_id")

write_csv(all,"01_r_processed_data/unnormalized_metabolomics.csv")

##### Normalizing the metabolomics data ######

standard = all %>% 
  filter(row_id == "RID_6043") %>% 
  select(metabolomics_id,standard_value = value)

#The normalization that Robert used was to divide the value by the standard value, multiply by 1E6 ( so it isn't negative) and then take the log10 of it. 
norm = all %>% 
  inner_join(standard, by = "metabolomics_id") %>% 
  mutate(norm_value = log10((value/standard_value) * 1E6))

inf = norm %>% 
  filter(norm_value == -Inf) %>% 
  select(row_id,value,standard_value,norm_value) %>% 
  pull(row_id) %>% 
  unique()

`%!in%` = Negate(`%in%`)

normalized_data = norm %>% 
  filter(row_id %!in% inf,
         row_id != "RID_6043")

write_csv(normalized_data,"01_r_processed_data/normalized_metabolomics.csv")
```

## A - Volcano plots of differentially abundant features.

```{r,fig.height=4, fig.width = 4}
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv",na = c("N/A","NA"))
## Reading in the normalized metabolomics data ##
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>%
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ infection_status) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Extracting the values for each feature for healthy patients
healthy = normalized_data %>% 
  filter(infection_status == "uninfected") %>% 
  group_by(row_id) %>% 
  summarise(healthy = mean(norm_value))

# Extracting the values for each feature for infected patients
infected = normalized_data %>% 
  filter(infection_status == "infected") %>% 
  group_by(row_id) %>% 
  summarise(infected = mean(norm_value))

## Combining the healthy and infected so we can caclulated the log2fc ##
log2fc = inner_join(healthy,infected,by = "row_id") %>% 
  mutate(log2fc_infected_healthy = log2(infected/healthy))

## Joining the statistics and log2 fold change ##  
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE))

f3a = vp %>% 
  ggplot(aes(log2fc_infected_healthy,-log10(p.adj_fdr),color = annotated))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("log2FC (Infected/Healthy)")+
  ggprism::theme_prism(base_size = 10)+
  theme(legend.position = "none")

f3a

ggsave(f3a,filename = "02_figures/f3_infection_v_healthy_metabolomics/f3a.pdf",height = 4, width = 4)
```

## B - breakdown on number of features ID

```{r}
## Reading in the data ##
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv",na = c("N/A","NA"))

normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

unnormalized_data = read_csv("01_r_processed_data/unnormalized_metabolomics.csv")

## Calculating the number of features ##
n_identified_features_gnps = nrow(gnps_annotations)
total_features = length(unique(unnormalized_data$row_id))
features_after_norm = length(unique(normalized_data$row_id))

percent_of_total_features_id = round(n_identified_features_gnps / total_features * 100,)

percent_of_norm_features_id = round(n_identified_features_gnps / features_after_norm * 100,0)

# Putting together a data frame # 
feature_count = tibble(`# total features` = total_features,
                       `# features after normalization` = features_after_norm,
                       `# of identified features by GNPS` = n_identified_features_gnps,
                       `% of total features ided` = percent_of_total_features_id,
                       `% of normalized features ided` = percent_of_norm_features_id) %>% 
  pivot_longer(1:length(.))

#Making a plot of the table
pdf("02_figures/f3_infection_v_healthy_metabolomics/f3b.pdf",height = 4, width = 4)
gridExtra::grid.table(feature_count)
dev.off()
```

## C - enrichment analysis of Metabolites in Infected Samples

```{r,fig.height = 3, fig.width = 4}
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

## Structuring the data for enrichment Analysis ##
gnps_enrichment_data = gnps_annotations %>% 
  #Manual conversion of unreasonable name to what it is listed as in MASSBANK
  select(compound_name_cleaned,npclassifier_superclass,npclassifier_class,npclassifier_pathway) %>% 
  mutate(GO_id = paste0(npclassifier_superclass,"_",npclassifier_class,"_",npclassifier_pathway))

#Making a makeshift "GO" type database using the GNPS annotations. 
TERM2GENE = gnps_enrichment_data %>% 
  dplyr::select(TERM = GO_id,GENE = compound_name_cleaned) %>% 
  as.data.frame()

TERM2NAME = gnps_enrichment_data %>% 
  select(-compound_name_cleaned) %>% 
  pivot_longer(cols = 1:3) %>% 
  select(keys = GO_id,columns = value) %>% 
  as.data.frame()

## Extracting the row_ids that were significant from the volcano plot ##
stat_different = vp %>% 
  filter(p.adj_fdr <= 0.05)

#
#converting RID to compound_name
stat_different_map = inner_join(vp,gnps_annotations,by = "row_id")

## Extracting the row ids that were upregulated in 
up = stat_different_map %>% 
  filter(log2fc_infected_healthy > 0 ) %>% 
  pull(compound_name_cleaned)

# Performing an enrichment
enriched_in_infected = clusterProfiler::enricher(up,
         universe = gnps_enrichment_data$compound_name_cleaned,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME
         )

f3c = clusterProfiler::cnetplot(enriched_in_infected,cex_label_category = 1,cex_label_gene = 0.5 )+
  theme(legend.position = "none")

f3c

ggsave("02_figures/f3_infection_v_healthy_metabolomics/f3c.pdf",f3c,height = 3,width = 4)
```

```{r}
down = stat_different_map %>% 
  filter(log2fc_infected_healthy < 0 ) %>% 
  pull(compound_name_cleaned)

# Performing an enrichment
enriched_in_healthy = clusterProfiler::enricher(down,
         universe = gnps_enrichment_data$compound_name_cleaned,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,pAdjustMethod = "none"
         )

enriched_in_healthy

```

^ There are no enriched terms here.

### Staph data for supplement

```{r,fig.height = 4,fig.height = 4}
staph_metabolomics = readxl::read_excel("00_r_input_data/staph_metabolomics.xlsx",skip = 1)

## Structuring the data for enrichment Analysis ##
gnps_enrichment_data = staph_metabolomics %>% 
  #Manual conversion of unreasonable name to what it is listed as in MASSBANK
  select(Unique_ID,Superclass,Class,Subclass) %>% 
  mutate(GO_id = paste0(Superclass,"_",Class,"_",Subclass))

#Making a makeshift "GO" type database using the GNPS annotations. 
TERM2GENE = gnps_enrichment_data %>% 
  dplyr::select(TERM = GO_id,GENE = Unique_ID) %>% 
  as.data.frame()

TERM2NAME = gnps_enrichment_data %>% 
  select(-Unique_ID) %>% 
  pivot_longer(cols = 1:3) %>% 
  select(keys = GO_id,columns = value) %>% 
  as.data.frame()

## Extracting the row_ids that were significant from the volcano plot ##
stat_different = staph_metabolomics %>% 
  mutate(Infection_pvalue = as.numeric(Infection_pvalue)) %>% 
  filter(Infection_pvalue <= 0.05)

## Extracting the row ids that were upregulated in 
up = stat_different %>% 
  filter(Infection_FC > 1) %>% 
  pull(Unique_ID)

# Performing an enrichment
enriched_in_infected = clusterProfiler::enricher(up,
         universe = gnps_enrichment_data$Unique_ID,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME
         )

## Here is the data frame with only the identified unique IDs
names = staph_metabolomics %>% 
  filter(Unique_ID %in% c(2510,2915,1917,2800,3125,3138,2263,2240,2470,3507,2767,1898,3509,1944))

# Manually converting the RID to reasonable names
fixed_names_vector = c("2510",
  "12-Ketodeoxycholic acid",
  "2-((4R)-4-((3R,5S,7R,9S,10S,13R,14S,17R)-3,7-dihydroxy-10,13-dimethylhexadecahydro-1H-cyclopenta[a]phenanthren-17-yl)pentanamido)ethane-1-sulfonic acid",
  "2-((4R)-4-((3R,5S,7R,9S,10S,13R,14S,17R)-3,7-dihydroxy-10,13-dimethylhexadecahydro-1H-cyclopenta[a]phenanthren-17-yl)pentanamido)ethane-1-sulfonic acid",
  "(2S,6R)-2-methyl-6-((3R,5S,7R,9S,10S,12S,13R,14S,17R)-3,7,12-trihydroxy-10,13-dimethylhexadecahydro-1H-cyclopenta[a]phenanthren-17-yl)heptanoic acid",
  "3138",
  "2-((4R)-4-((3R,5S,7R,9S,10S,13R,14S,17R)-3,7-dihydroxy-10,13-dimethylhexadecahydro-1H-cyclopenta[a]phenanthren-17-yl)pentanamido)ethane-1-sulfonic acid",
  "Tauroursodeoxycholic acid",
  "3beta-Hydroxy-5-cholenoic acid",
  "Taurocholic acid",
  "2767",
  "Taurocholic acid",
  "Taurocholic acid",
  "Glycocholic acid")

fixed_names = paste0(fixed_names_vector,collapse = "/")

enriched_in_infected@result$geneID = fixed_names

up_plot = clusterProfiler::cnetplot(enriched_in_infected,cex_label_category = 1, cex_label_gene = 0.4)+
  theme(legend.position = "none")

up_plot

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_in_staph_metabolomics.pdf",up_plot,width = 4, height = 4)
```

```{r,fig.height = 4, fig.width = 4}
## Extracting the row ids that were upregulated in 
down = stat_different %>% 
  filter(Infection_FC < 1) %>% 
  pull(Unique_ID)

# Performing an enrichment\

# Performing an enrichment
enriched_in_healthy = clusterProfiler::enricher(down,
         universe = gnps_enrichment_data$Unique_ID,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME
         )

## Here is the data frame with only the identified unique IDs
names = staph_metabolomics %>% 
  filter(Unique_ID %in% c(554,132,450,1939,33,372,99,557,1057,586,434,128))

fixed_names_vector = c("1-Myristoyl-sn-glycero-3-phosphocholine",
                "132",
                "1-Octadecyl-2-acetyl-sn-glycero-3-phosphocholine",
                "1-Myristoyl-sn-glycero-3-phosphocholine",
                "1-Pentadecanoyl-sn-glycero-3-phosphocholine",
                "372",
                "1-Octadecyl-sn-glycero-3-phosphocholine",
                "1,2-Dioctanoyl PC",
                "1-(1Z-Octadecenyl)-sn-glycero-3-phosphocholine",
                "1-(1Z-Octadecenyl)-sn-glycero-3-phosphocholine",
                "434",
                "1-O-Hexadecyl-2-O-(2E-butenoyl)-sn-glyceryl-3-phosphocholine")

fixed_names = paste0(fixed_names_vector,collapse = "/")

enriched_in_healthy@result$geneID = fixed_names

healthy_plot = clusterProfiler::cnetplot(enriched_in_healthy,cex_label_category = 1,show_category = 100,cex_label_gene = 0.5)+
  theme(legend.position = "none")
healthy_plot

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_in_staph_metabolomics.pdf",healthy_plot,width = 4, height = 4)

```

## D - Assesment of top biomarkers

### Data processing
```{r,fig.height = 8, fig.width = 8}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

## Switching to the wide format ### 
norm_wide = normalized_data %>%
  select(-row_m_z,-row_retention_time,-value) %>% 
  tidyr::pivot_wider(names_from = row_id,values_from = norm_value)

## Looking for only columns that were annotated ## 
gnps_and_in_norm = intersect(unique(gnps_annotations$row_id), colnames(norm_wide))

gnps_annotated_metabolites = norm_wide %>% 
  as.data.frame() %>% 
  select(condition,gnps_and_in_norm)

## Extracting the numeric data ##
efs_sel = gnps_annotated_metabolites %>% 
  mutate(classlabels = case_when(condition %in% c("faecalis","faecium") ~ 0,
                           TRUE ~ 1)) %>% 
  select(classlabels,2:length(.)) %>% 
   #absolutely needs to be a data frame and not a tibble for EFS package
  as.data.frame()

# Seting seed incase this isn't done internally in EFS
set.seed(2)

## Running EFS feature selection ##
ensemble_result = EFS::ensemble_fs(efs_sel,
                                   classnumber = 1,cor_threshold = 0)

## Formating the data as long ##
 long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "row_id")

## Extracting top metabolites ##
top = long %>% 
  group_by(row_id) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

## Pulling top N metabolites ##
topn = top[1:10,] %>% 
  pull(row_id)

#writing file contining the topn row ids 
topn_metabolomics = tibble(row_id = topn)

write_csv(topn_metabolomics,"01_r_processed_data/top10_infected_v_healthy_metabolomics_biomarkers.csv")

## Filtering the data, combining annotation names ##
filtered = long %>% 
  filter(row_id %in% topn) %>% 
  inner_join(gnps_annotations,by = "row_id")

## Plotting the data ## 
# f3d = filtered %>% 
#   ggplot(aes(reorder(compound_name_cleaned,value),value,fill = model))+
#   geom_col()+
#   coord_flip()+
#   viridis::scale_fill_viridis(discrete = TRUE)+
#   ggtitle("Top 10 Performing Metabolite Biomarkers of Infection")+
#   xlab("Metabolite Biomarkers")
# 
# f3d

#ggsave("02_figures/f3_infection_v_healthy_metabolomics/f3d.pdf",f3d,height = 8,width = 8,units = "in")
```

### Plotting

```{r,fig.height = 8, fig.width = 8}
## Arranging Data so it is in the correct order ### 
ROC_data = dplyr::filter(normalized_data,row_id %in% topn[1:2]) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = c("row_id" = "row_id")) %>% 
mutate(condition = case_when(condition == "healthy" ~ "healthy",
                             TRUE ~ "infected"))

source("functions.R")

top = make_violin_plot_metabolomics(ROC_data,topn = topn[1:2],ncol = 2,pname = "p.signif")

bottom = make_roc_metabolomics(ROC_data,ncol = 2)

f3d = ggpubr::ggarrange(top,bottom,nrow = 2)

f3d

ggsave("02_figures/f3_infection_v_healthy_metabolomics/f3d.pdf",f3d,height=8,width=8)
```

# Figure 4- Proteomic Differences between Enterococcus Faecium and Faecalis

## A- Volcano plot

```{r,fig.height = 4, fig.width =4}
faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid" = "ProteinID"))

  f4a = faecalis_v_faecium %>% 
  ggplot(aes(log2FC,-log10(adj.pvalue)))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("Log2FC Faecalis / Faecium")+
  ggprism::theme_prism(base_size = 10)

f4a

ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4a.pdf",f4a,height = 4, width = 4)
```

## B Network analysis up in Faecalis

```{r,fig.height = 4, fig.width = 8}
up_in_faecalis = faecalis_v_faecium %>% 
  filter(log2FC>0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(Gene)

#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
    # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

enrich = clusterProfiler::enricher(up_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

f4b = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                cex_label_gene = 0.5,
                                showCategory = 100)+
  theme(legend.position = "none")

f4b

ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4b.pdf",height = 4, width =8)
```

## C - Immunoglobulin comparisons

```{r,fig.height = 3.8, fig.width = 4}
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

anot = read_csv("00_r_input_data/all_protein_annotations.csv")

sum_annotated = left_join(protein_sum,anot,by = c("uniprotid" = "Entry"))

ig = sum_annotated %>% 
  filter(grepl(".*Immunoglobulin.*",Protein.names)) %>% 
  filter(!grepl(".*receptor.*",Protein.names)) %>% 
  #Removing one problematic immunglobulin. Negative values for 3 of the samples messed things up. 
  #Doesn't change the story, just the plot. 
  filter(uniprotid != "A0A0B4J2B5")

stat = ig %>% 
  rstatix::t_test(Abundance ~ Condition)

f4c = ig %>% 
  ggplot(aes(Condition,Abundance),size = 0.5)+
  geom_violin()+
  geom_boxplot()+
  ggpubr::stat_pvalue_manual(stat,
                             y.position = 22,
                             step.increase = 0.1)+
  ggprism::theme_prism(base_size = 10)

f4c

ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4c.pdf",f4c,height = 3.8,width = 4)
```

## D - terms up in faecium.

```{r,fig.height = 3.8, fig.width = 4}
up_in_faecium = faecalis_v_faecium %>% 
  filter(log2FC<0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(Gene)

enrich = clusterProfiler::enricher(up_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

f4d = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                cex_label_gene = 0.5,
                                showCategory = 100)+
  theme(legend.position = "none")

f4d

ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4d.pdf",f4d,height = 3.8, width = 4)
```

## E - Assesment of Biomarkers

### Data Processing

```{r,fig.height=8,fig.width = 8}
library(EFS)
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

clean = pl %>% 
  dplyr::group_by(Mixture,Gene,Channel,Condition) %>% 
  dplyr::summarise(sum = sum(Abundance))

faecium_v_faecalis_l = clean %>% 
  filter(Condition %in% c("Faecium", "Faecalis")) %>% 
  mutate(classlabels = case_when(Condition == "Faecalis" ~ 1,
                                   TRUE ~ 0)) %>%
  # EFS can't handle the - charachter. 
  mutate(Gene = gsub("-","_",Gene))

faecium_v_faecalis_w = faecium_v_faecalis_l %>% 
  pivot_wider(names_from = Gene,
              values_from = sum) %>% 
  mutate(sampleid = paste0(Mixture,".",Channel)) %>% 
  ungroup() %>% 
  dplyr::select(-Mixture,-Channel,-Condition,-sampleid) %>% 
  select_if(~ !any(is.na(.))) %>% 
  as.data.frame()

#need to come up with some sort of way to combine all of the fractions and plexes
# The problem is that na.omit filters everything out. 
ensemble_result = EFS::ensemble_fs(faecium_v_faecalis_w,1,cor_threshold = 0)

long = ensemble_result %>% 
  as.data.frame() %>% 
  #Converting gene names back
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "Gene") %>% 
  mutate(Gene = gsub("_","-",Gene))

top = long %>% 
  group_by(Gene) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

topn = top[1:10,] %>% 
  pull(Gene)

# writing to file 
topn_proteomics = tibble(Gene = topn)

write_csv(topn_proteomics,"01_r_processed_data/top10_faecalis_v_faecium_proteomics_biomarkers.csv")

# f4d = long %>% 
#   filter(Gene %in% topn) %>% 
#   ggplot(aes(reorder(Gene,value),value,fill = model))+
#   geom_col()+
#   coord_flip()+
#   viridis::scale_fill_viridis(discrete = TRUE)+
#   ggtitle("Top 10 Biomarkers Differentiating Faecalis from Faecium by EFS")+
#   xlab("Protein Biomarkers")
# 
# f4d

# ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4d.pdf",f4d,height = 8,width = 8)
```

### Plotting

```{r,fig.height=4,fig.width=8}

ROC_data = dplyr::filter(protein_sum,Gene %in% topn[1:2]) %>% 
  dplyr::filter(Condition %in% c("Faecalis","Faecium"))

source("functions.R")

f4d = make_violin_plot_proteomics(ROC_data,topn[1:2],ncol = 2,pname = "p.signif")
f4e = make_roc_proteomics(ROC_data,topn[1:2],ncol=2)

f4d = ggpubr::ggarrange(f4d,f4e,nrow = 2)

ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4d.pdf",f4d,height=8,width=8)
```

# Figure 5 - Metabolomic Differences between Faecalis and Faecium Infected samples

## A - Volcano plot of data.

```{r,fig.height = 3.8, fig.width = 4}
## Reading in Data
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>%
    filter(condition %in% c("faecalis","faecium")) %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ condition) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,condition,norm_value)

# Extracting just the faecalis data
faecalis = log2fc_data %>% 
  filter(condition == "faecalis") %>% 
  group_by(row_id) %>% 
  summarise(faecalis = mean(norm_value))

# Extracting just the faecium data 
faecium = log2fc_data %>% 
  filter(condition == "faecium") %>% 
  group_by(row_id) %>% 
  summarise(faecium = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(faecalis,faecium,by = "row_id") %>% 
  mutate(log2fc_faecalis_faecium = log2(faecalis/faecium))

# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE))

annotations = vp %>% 
  filter(p.adj_fdr <= 0.05,
         annotated == TRUE) %>% 
  inner_join(gnps_annotations, by = "row_id")

f5a = vp %>% 
  ggplot(aes(log2fc_faecalis_faecium,-log10(p.adj_fdr),color = annotated))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("log2FC (Faecalis/Faecium)")+
  ggrepel::geom_label_repel(data = annotations,aes(label = compound_name_cleaned,x= log2fc_faecalis_faecium, y = -log10(p.adj_fdr)),size = 2.5,label.size = 0.1)+
  ggprism::theme_prism(base_size = 10)+
  theme(legend.position = "none")

f5a

ggsave("02_figures/f5_faecalis_v_faecium_metabolomics/f5a.pdf",f5a,height = 3.8, width =4)
```

## B - Biomarker Assesment

### Data Processing
```{r,fig.height = 8 , fig.width = 8}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv") %>% 
#Removing reduntant RIDs. These were found to be the same as other top biomarkers
  filter(row_id != "RID_8856",
         row_id != "RID_22529",
         row_id != "RID_22270",
         row_id != "RID_9436")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

## Switching to the wide format ### 
norm_wide = normalized_data %>%
  select(-row_m_z,-row_retention_time,-value) %>% 
  tidyr::pivot_wider(names_from = row_id,values_from = norm_value)

## Looking for only columns that were annotated ## 
gnps_and_in_norm = intersect(unique(gnps_annotations$row_id), colnames(norm_wide))

gnps_annotated_metabolites = norm_wide %>% 
  as.data.frame() %>% 
  select(condition,gnps_and_in_norm)

## Extracting the numeric data ##
efs_sel = gnps_annotated_metabolites %>% 
  filter(condition %in% c("faecalis","faecium")) %>% 
  mutate(classlabels = case_when(condition == "faecalis" ~ 0,
                           TRUE ~ 1)) %>% 
  select(classlabels,2:length(.)) %>% 
   #absolutely needs to be a data frame and not a tibble for EFS package
  as.data.frame()

# Seting seed incase this isn't done internally in EFS
set.seed(2)

## Running EFS feature selection ##
ensemble_result = EFS::ensemble_fs(efs_sel,
                                   classnumber = 1,
                                   cor_threshold = 0)

## Formating the data as long ##
 long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "row_id") %>% 
  inner_join(gnps_annotations,by = "row_id")

## Extracting top metabolites ##
top = long %>% 
  group_by(row_id,compound_name_cleaned) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

## Pulling top N metabolites ##
topn = top[1:10,] %>% 
  pull(row_id)

topn_metabolomics = tibble(row_id = topn)

write_csv(topn_metabolomics,"01_r_processed_data/top10_faecalis_v_faecium_metabolomics_biomarkers.csv")

## Filtering the data, combining annotation names ##
filtered = long %>% 
  filter(row_id %in% topn)

## Plotting the data ## 
EFS_plot = filtered %>% 
  ggplot(aes(reorder(compound_name_cleaned,value),value,fill = model))+
  geom_col()+
  coord_flip()+
  viridis::scale_fill_viridis(discrete = TRUE)+
  xlab("Metabolite Biomarkers")+
  theme(legend.position = "bottom")

EFS_plot

#ggsave("02_figures/f5_faecalis_v_faecium_metabolomics/f5b.pdf",f5b,height = 8,width = 8)
```

mycophenolic acid and pc(17:1/0:0 + c25h51n1o7p1) were detected as two seperate row IDs. They were both found to be among the best biomarkers independantly from each other.

### Plotting

```{r,fig.height = 8, fig.width = 8}
ROC_data = dplyr::filter(normalized_data,row_id %in% topn[1:2]) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  filter(condition != "healthy") %>% 
  mutate(infection_status = condition)

top = make_violin_plot_metabolomics(ROC_data,topn[1:2],ncol = 2,pname = "p.signif")
bottom = make_roc_metabolomics(ROC_data,ncol = 2)

f5b = ggpubr::ggarrange(top,bottom,nrow = 2)

f5b

ggsave("02_figures/f5_faecalis_v_faecium_metabolomics/f5b.pdf",f5b,height = 8, width = 8)
```

# Figure 6 - Prediction of Clinical Outcome

## A- Unsupervised clustering of Proteomics and Metabolomic data combined

```{r,fig.height = 8, fig.width = 12}
##Proteomics data ##
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  select(tmt_id = sample_id,everything())

## Clinical metadata ##
map = readxl::read_excel("00_r_input_data/proteomics_sample_mapping.xlsx") %>% 
  mutate(tmt_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(tmt_id,sample_id = sample_name)

md = read_csv("00_r_input_data/clinical_metadata.csv")

md_final = inner_join(map,md,by = "sample_id")

clean_clinical = inner_join(pl,md_final, by = "tmt_id")

clean_clinical_wide = clean_clinical %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #removing the samples without death during admission info
  filter(!is.na(death_during_admission)) %>% 
  #filter(sample_id.y != "S49",
         #sample_id.y != "S76") %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame()

### Combining all data ### 
all= clean_clinical_wide

# Getting into a format where we can do the clustering ## 
clust_data = all[,3:283] %>% 
   #Remove columns that only contain 0s 
  select_if(colSums(.) != 0) %>% 
  as.data.frame()

annotation = all[,2] %>% 
  as.data.frame()

## Fixing the row names so they match for the heatmap ##
rownames(annotation) = all$sample_id

rownames(clust_data) = all$sample_id

#calculating distance matrix
d = dist(clust_data)
## Generating a dendrogram without a heatmap ##
library(dendextend)

dend = as.dendrogram(hclust(d, method = "ward.D2"))

# let's add some color: Red = dead, black = okay
colors_to_use = c("#000000","#FF0000")[as.numeric(as.factor(all$death_during_admission))]

#getting in the same order as dendrogram
colors_to_use = colors_to_use[order.dendrogram(dend)]

# Now we can color the brancehes
dend = branches_color(dend,col = colors_to_use,) %>% 
  set("branches_lwd", 8)

# Plotting for manuscript
pdf("02_figures/f6_prediction_of_clinical_outcome_proteomics/f6a.pdf",height = 8,
    width = 12)
plot(dend)
dend %>% rect.dendrogram(k=2, 
                           border = 8, lty = 5, lwd = 2)
dev.off()

#Plotting for this document
plot(dend)
dend %>% rect.dendrogram(k=2, 
                           border = 8, lty = 5, lwd = 2)

# Checking stats to see if there are significant differences between clusters when k=2
xtab <- as.table(rbind(c(4, 8), c(13, 53)))
dimnames(xtab) <- list(
  group = c("k1", "k2"),
  death = c("yes", "no")
)

fisher_test(xtab)
```

## B Volcano plot of Proteomics

```{r,eval = FALSE}
### Need to rerun msstats, this time changing the condition to be death upon admission ###
# read in clincal metadata

cmd = read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  select(BioReplicate = sample_id,death_during_admission)

# Read MSstats.csv file.
data = read_csv("00_r_input_data/msstats.csv")
#Note that the condition of the bridge channel needs to be labeled as Norm
annotation = read_csv("00_r_input_data/MSstatsTMT_annotation.csv") %>% 
  inner_join(cmd,by = "BioReplicate") %>% 
  select(-Condition) %>% 
  rename(Condition = death_during_admission) %>% 
  mutate(Condition = as.factor(as.character(Condition))) %>% 
  mutate(Condition = case_when(is.na(Condition) ~ "unknown",
                               TRUE ~ Condition))

mstats = MSstatsTMT::PhilosophertoMSstatsTMTFormat(input = data,annotation)

# use MSstats for protein summarization
quant.msstats = MSstatsTMT::proteinSummarization(mstats,
                                      method="msstats",
                                      global_norm=TRUE,
                                      reference_norm=TRUE,
                                      remove_norm_channel = TRUE,
                                      remove_empty_channel = TRUE)

# Taking the pl_data and formatting appropriately
pl_data = quant.msstats$ProteinLevelData %>% 
  tidyr::separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>%
  # Converting to gene name
  inner_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  mutate(sample_id = paste0(Mixture,".",Channel))

write_csv(pl_data,"01_r_processed_data/quant_msstats_protein_level_data_death.csv")

### Preforming the statistics ### 
test.pairwise = MSstatsTMT::groupComparisonTMT(quant.msstats,
                                               moderated = TRUE)

stats =test.pairwise$ComparisonResult %>% 
  filter(Label == "FALSE vs TRUE") %>% 
  separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>% 
  inner_join(protein_to_gene,by = c("uniprotid" = "ProteinID"))

write_csv(stats,"01_r_processed_data/death_vs_life.csv")
```

```{r,fig.height=3.8, fig.width = 4}

dvl = read_csv("01_r_processed_data/death_vs_life.csv") %>% 
  filter(is.na(issue))

f6a = dvl %>% 
  ggplot(aes(log2FC,-log10(adj.pvalue)))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("Log2FC (Live/Dead)")+
  ggprism::theme_prism(base_size = 10)

f6a

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6a.pdf",f6a,height = 3.8, width = 4)

# Extracting the signficiant proteins from the volcano plot for later use. 
sig = dvl %>% 
  filter(adj.pvalue <= 0.05)

up_in_live = sig %>% 
  filter(log2FC>0) %>% 
  pull(Gene)

up_in_dead = sig %>% 
  filter(log2FC<0) %>% 
  pull(Gene)
```

## B - GO Enrichment in Dead

```{r,fig.height = 4, fig.width = 4}
#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_dead,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

f6b = clusterProfiler::cnetplot(enrich, categorySize="pvalue",circular = FALSE,categorySize="pvalue",cex_label_category = 1,showCategory = 100,cex_label_gene = 0.5)+
  theme(legend.position = "none")

f6b

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6b.pdf",f6b,height = 4, width = 4)

```

## D - GO Enrichment in Live

```{r,fig.height = 4, fig.width = 4}
#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

TERM2NAME = AnnotationDbi::select(GO.db::GO.db,keys = unique(TERM2GENE$TERM),columns = c("TERM")) %>% 
  as.data.frame()

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_live,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

f6c = clusterProfiler::cnetplot(enrich, categorySize="pvalue",
                              circular = FALSE,
                              categorySize="pvalue",
                              cex_label_category = 1,
                              showCategory = 100,
                              cex_label_gene = 0.5, )+
  theme(legend.position = "none")
f6c

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6c.pdf",f6c)
```

## D - Proteomic Biomarkers to predict death

### Data Processing
```{r,fig.height = 6, fig.width = 16}
##reading in the data##
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  rename(tmt_id = sample_id)

## Clinical metadata ##
map = readxl::read_excel("00_r_input_data/proteomics_sample_mapping.xlsx") %>%
  mutate(tmt_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(tmt_id,sample_id = sample_name)

md = read_csv("00_r_input_data/clinical_metadata.csv")

md_final = inner_join(map,md,by = "sample_id")

clean_clinical = inner_join(pl,md_final, by = "tmt_id") %>% 
  # Don't need the healthy samples here
  filter(Condition != "Healthy") %>%
  #IGKC gets identified by EFS even though it is a terrible biomarker. Not sure why, just going to remove it
  filter(Gene != "IGKC") %>% 
   # - messes up EFS for some reason, very annoying. 
  mutate(Gene = gsub("-","_",Gene)) %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #Removing samples with NA.
  filter(!is.na(death_during_admission)) %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame() %>% 
  select(-sample_id)

set.seed(2)
ensemble_result = EFS::ensemble_fs(clean_clinical,1,cor_threshold = 0)

long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "Gene") %>% 
  #converting genes back to proper notation. 
  mutate(Gene = gsub("_","-",Gene))

top = long %>% 
  group_by(Gene) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

topn = top[1:10,] %>% 
  pull(Gene)

topn_out = tibble(Gene = topn)

write_csv(topn_out,"01_r_processed_data/top10_live_v_dead_proteomics_biomarkers.csv")

filtered = long %>% 
  filter(Gene %in% topn)

EFS = filtered %>% 
  ggplot(aes(reorder(Gene,value),value,fill = model))+
  geom_col()+
  coord_flip()+
  viridis::scale_fill_viridis(discrete = TRUE)+
  ggtitle("Top 10 Performing Protein Biomarkers")+
  xlab("Protein Biomarkers")

EFS
```

### Plotting

```{r,fig.height=8,fig.width=8}
source("functions.R")

data = inner_join(pl,md_final, by = "tmt_id") %>%
  filter(Condition != "Healthy") %>% 
  mutate(Condition = death_during_admission) %>% 
  filter(!is.na(Condition))

top = make_violin_plot_proteomics(data,topn[1:2],pname = "p.signif",ncol = 2)

bottom = make_roc_proteomics(data,topn[1:2],ncol = 2)

f6d = ggpubr::ggarrange(top,bottom,nrow = 2)

f6d

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6d.pdf",f6d,height = 8, width = 8)
```
## E- Metabolomics Volcano Plot

```{r,fig.height = 3.8, fig.width = 4 }
## Reading in Data
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv") %>% 
  filter(!is.na(death_during_admission))

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ death_during_admission) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,death_during_admission,norm_value)

# Extracting just the faecalis data
dead = log2fc_data %>% 
  filter(death_during_admission == TRUE) %>% 
  group_by(row_id) %>% 
  summarise(dead = mean(norm_value))

# Extracting just the faecium data 
live = log2fc_data %>% 
  filter(death_during_admission == FALSE) %>% 
  group_by(row_id) %>% 
  summarise(live = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(dead,live,by = "row_id") %>% 
  mutate(log2fc_live_dead = log2(live/dead))

# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE))

annotations = vp %>% 
  filter(p.adj_fdr <= 0.05,
         annotated == TRUE) %>% 
  inner_join(gnps_annotations, by = "row_id")

f6e = vp %>% 
  ggplot(aes(log2fc_live_dead,-log10(p.adj_fdr),color = annotated))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("log2FC (Live/Dead)")+
  theme(legend.position = "bottom")+
  ggrepel::geom_label_repel(data = annotations,aes(label = compound_name_cleaned,x= log2fc_live_dead, y = -log10(p.adj_fdr)),size = 2.5,label.size = 0.1,nudge_y = 0.1)+
  ggprism::theme_prism(base_size = 10)+
  theme(legend.position = "none")

f6e

ggsave(filename = "02_figures/f6_prediction_of_clinical_outcome/f6e.pdf",f6e,height = 3.8, width = 4)
```

## F- Evaluation of Metabolomics Biomarkers to Predict Death

```{r,fig.height = 14,fig.width = 14}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

## Switching to the wide format ### 
norm_wide = normalized_data %>%
  select(-row_m_z,-row_retention_time,-value) %>% 
  tidyr::pivot_wider(names_from = row_id,values_from = norm_value)

## Looking for only columns that were annotated ## 
gnps_and_in_norm = intersect(unique(gnps_annotations$row_id), colnames(norm_wide))

gnps_annotated_metabolites = norm_wide %>% 
  as.data.frame() %>% 
  select(death_during_admission,gnps_and_in_norm) %>% 
  filter(!is.na(death_during_admission))

# Seting seed incase this isn't done internally in EFS
set.seed(2)

## Running EFS feature selection ##
ensemble_result = EFS::ensemble_fs(gnps_annotated_metabolites,
                                   classnumber = 1,cor_threshold = 0)

## Formating the data as long ##
 long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.),names_to = "row_id")

## Extracting top metabolites ##
top = long %>% 
  group_by(row_id) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum)

## Pulling top N metabolites ##
topn = top[1:10,] %>% 
  pull(row_id)

topn_out = tibble(row_id = topn )

write_csv(topn_out,"01_r_processed_data/top10_live_v_dead_metabolomics_biomarkers.csv")

data = normalized_data %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  filter(!is.na(death_during_admission)) %>% 
  filter(row_id %in% topn[1:2]) %>% 
  mutate(condition = as.character(death_during_admission))

top = make_violin_plot_metabolomics(data,ncol = 2,topn = topn[1:2],pname = "p.signif")
bottom = make_roc_metabolomics(data,ncol = 2)

f6f = ggpubr::ggarrange(top,bottom,nrow = 2)

f6f
ggsave("02_figures/f6_prediction_of_clinical_outcome/f6f.pdf",f6f,height = 8,width = 8)

```

# Suplemental Figures

## S1 - Overview of Analysis Workflow

This is generated in draw.io

## S2 - Networks

See Figure 2 B and Figure 3 C

## S3 - Known Biomarkers of Infection

### A - ROC curves.

```{r,fig.height=4, fig.width = 8}
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

known_indicatiors_of_bacterial_infection = c("CRP","SAA1")

f = dplyr::filter(pl,Gene %in% known_indicatiors_of_bacterial_infection) %>%
  mutate(Condition = case_when(Condition == "Healthy" ~ "Healthy",
                               TRUE ~ "Infected"))

data = f

  list = list()

  for(i in known_indicatiors_of_bacterial_infection){

    loop = filter(data, Gene == i)

    mod = glm(as.factor(Condition) ~ Abundance, 
              data = data,
              family = "binomial")

    predicted <- predict(mod ,family = "binomial", loop, type="response")

    #define object to plot
    rocobj <- pROC::roc(loop$Condition, predicted)

    auc = round(as.numeric(gsub(".*: ", "",rocobj$auc)),2)
    #create ROC plot
    p1 = pROC::ggroc(rocobj)+
      ggtitle(paste0(i," AUC: ",auc))

    list[[i]] = p1
  }
  s3a = do.call("grid.arrange", c(list, ncol = 2),)
  s3a

ggsave("03_supplemental/s3_known_biomarkers_of_infection/s3a.pdf",s3a)

```

### B - Violin Plot

```{r,fig.height = 6, fig.width = 8 }
#Loading and processing the data 
#reading in the data
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

# C-reactive Protein = "P02741
# Calcitonin = P01258
# Serum Amyloid A = P0DJI8

known_indicatiors_of_bacterial_infection = c("CRP","SAA1")

f = dplyr::filter(pl,Gene %in% known_indicatiors_of_bacterial_infection)

library(rstatix)
library(ggpubr)

ttest <-
  f %>%
  group_by(Gene) %>%
  rstatix::t_test(Abundance ~ Condition) %>%
  add_y_position()

s3b = f %>% 
  ggplot(aes(Condition,Abundance))+
  geom_violin()+
  geom_point()+
  ggpubr::stat_pvalue_manual(ttest)+
  ylab("Abundance")+
  facet_wrap(~Gene)

s3b

ggsave("03_supplemental/s3_known_biomarkers_of_infection/s3b.pdf",s3b)
```
## S4 - Top 10 identified biomarkers Healthy vs Infected

### A- proteomics violin plot

```{r,fig.height=6,fig.width=16}
# Reading in data
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

topn = read_csv("01_r_processed_data/top10_infected_v_healthy_proteomics_biomarkers.csv") %>% pull(Gene)

# Processing data 
ROC_data = dplyr::filter(protein_sum,Gene %in% topn) #%>% 
  #dplyr::mutate(Condition = case_when(Condition %in% c("Faecalis","Faecium") ~ "infected",
                                    #  TRUE ~ Condition))

source("functions.R")

s4a = make_violin_plot_proteomics(ROC_data,topn)

s4a

ggsave("03_supplemental/s4_biomarkers_healthy_v_infected/s4a.pdf",s4a, height = 6 ,width=16)
```

### B- Proteomics ROC

```{r,fig.height=6,fig.width = 16}
# Processing data 
ROC_data = dplyr::filter(protein_sum,Gene %in% topn) %>% 
  dplyr::mutate(Condition = case_when(Condition %in% c("Faecalis","Faecium") ~ "infected",
                                     TRUE ~ Condition))

  source("functions.R")

s4b = make_roc_proteomics(ROC_data,topn)

s4b

ggsave("03_supplemental/s4_biomarkers_healthy_v_infected/s4b.pdf",s4b, height = 6 ,width=16)

```

### C - Metabolomics Violin

```{r,fig.height=6,fig.width = 16}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

topn = read_csv("01_r_processed_data/top10_infected_v_healthy_metabolomics_biomarkers.csv") %>% pull(row_id)

ROC_data = dplyr::filter(normalized_data,row_id %in% topn) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = c("row_id" = "row_id"))

source("functions.R")

s4c = make_violin_plot_metabolomics(ROC_data,topn)

ggsave("03_supplemental/s4_biomarkers_healthy_v_infected/s4c.pdf",s4c,height = 6 ,width=16)

```
### D - Metabolomics ROC

```{r,fig.height=6,fig.width = 16}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

topn = read_csv("01_r_processed_data/top10_infected_v_healthy_metabolomics_biomarkers.csv") %>% pull(row_id)

ROC_data = dplyr::filter(normalized_data,row_id %in% topn) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = c("row_id" = "row_id")) %>% 
  dplyr::mutate(condition = case_when(condition %in% c("faecalis","faecium") ~ "infected",
                                     TRUE ~ condition))

source("functions.R")

s4d = make_roc_metabolomics(ROC_data,)

ggsave("03_supplemental/s4_biomarkers_healthy_v_infected/s4d.pdf",s4d,height = 6 ,width=16)

```

## S5 - Top 10 Faecalis vs Faecium

### A Proteomics violin

```{r,fig.height=6, fig.width = 16}
# Reading in data
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

topn = read_csv("01_r_processed_data/top10_faecalis_v_faecium_proteomics_biomarkers.csv") %>% pull(Gene)

# Processing data 
ROC_data = dplyr::filter(protein_sum,Gene %in% topn)

source("functions.R")

s5a = make_violin_plot_proteomics(ROC_data,topn)

s5a

ggsave("03_supplemental/s5_biomarkers_faecalis_v_faecium/s5a.pdf",s5a,height = 6 ,width=16)
```
### B- Proteomics ROC

```{r,fig.height =6, fig.width = 16}
# Processing data 
ROC_data = dplyr::filter(protein_sum,Gene %in% topn) %>% 
  filter(Condition != "Healthy")

source("functions.R")

s5b = make_roc_proteomics(ROC_data,topn)

s5b

ggsave("03_supplemental/s5_biomarkers_faecalis_v_faecium/s5b.pdf",s5b,height = 6 ,width=16)
```

### C- metabolomics violin

```{r,fig.height = 6, fig.width = 16}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

topn = read_csv("01_r_processed_data/top10_faecalis_v_faecium_metabolomics_biomarkers.csv") %>% pull(row_id)

ROC_data = dplyr::filter(normalized_data,row_id %in% topn) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = c("row_id" = "row_id"))

s5c = make_violin_plot_metabolomics(ROC_data,topn)

ggsave("03_supplemental/s5_biomarkers_faecalis_v_faecium/s5c.pdf",s5c,height = 6 ,width=16)

```
### D- metabolomics ROC

```{r,fig.height =6, fig.width = 16}
ROC_data = dplyr::filter(normalized_data,row_id %in% topn) %>% 
  dplyr::arrange(factor(row_id,topn)) %>% 
  inner_join(gnps_annotations,by = c("row_id" = "row_id")) %>% 
  filter(condition != "healthy")

s5d = make_roc_metabolomics(ROC_data)

ggsave("03_supplemental/s5_biomarkers_faecalis_v_faecium/s5d.pdf",s5d,height = 6 ,width=16)
```

## S6- Top 10 Mortality vs Survival

### A- Proteomics violin

```{r, fig.height = 6, fig.width = 16}
##reading in the data##
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  rename(tmt_id = sample_id)

## Clinical metadata ##
map = readxl::read_excel("00_r_input_data/proteomics_sample_mapping.xlsx") %>%
  mutate(tmt_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(tmt_id,sample_id = sample_name)

md = read_csv("00_r_input_data/clinical_metadata.csv")

md_final = inner_join(map,md,by = "sample_id")

clean_clinical = inner_join(pl,md_final, by = "tmt_id") %>% 
  # Don't need the healthy samples here
  filter(Condition != "Healthy") %>% 
   # - messes up EFS for some reason, very annoying. 
  mutate(Gene = gsub("-","_",Gene)) %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #Removing samples with NA.
  filter(!is.na(death_during_admission)) %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame() %>% 
  select(-sample_id)

topn = read_csv("01_r_processed_data/top10_live_v_dead_proteomics_biomarkers.csv") %>% 
  pull(Gene)

data = inner_join(pl,md_final, by = "tmt_id") %>%
  filter(Condition != "Healthy") %>% 
  mutate(Condition = death_during_admission) %>% 
  filter(!is.na(Condition))

s6a = make_violin_plot_proteomics(data,topn,pname="p.signif")

ggsave("03_supplemental/s6_biomarkers_survival_v_mortality/s6a.pdf",s6a,height = 6, width = 16)
```

### B- Proteomics ROC

```{r,fig.height = 6, fig.width = 16}

s6b = make_roc_proteomics(data,topn)

ggsave("03_supplemental/s6_biomarkers_survival_v_mortality/s6b.pdf",s6b,height = 6, width = 16)

```

### C - Metabolomics violin

```{r,fig.height = 6, fig.width = 16}
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

topn = read_csv("01_r_processed_data/top10_live_v_dead_metabolomics_biomarkers.csv") %>% 
  pull(row_id)

data = normalized_data %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  filter(!is.na(death_during_admission)) %>% 
  filter(row_id %in% topn) %>% 
  mutate(condition = as.character(death_during_admission))

s6c = make_violin_plot_metabolomics(data,topn,ncol = 5,pname="p.signif")

ggsave("03_supplemental/s6_biomarkers_survival_v_mortality/s6c.pdf",s6c,height = 6,width = 16)

```

### D - Metabolomics ROC

```{r,fig.height = 6, fig.width = 16}
s6d = make_roc_metabolomics(data)

ggsave("03_supplemental/s6_biomarkers_survival_v_mortality/s6d.pdf",s6d,height = 6,width = 16)
```

## S7- metadata assesment of all biomarkers ided in the study

### A- Faecalis vs Faecium Proteomics.

```{r,fig.height = 8, fig.width = 16}

clinical_md = read_csv("00_r_input_data/clinical_metadata.csv")
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

# Defining the input metadata. Need to chose which columns to keep here. 
md_assesment_data= protein_sum %>% 
  filter(Condition != "Healthy") %>% 
  inner_join(clinical_md, by = c("BioReplicate" = "sample_id")) %>% 
  select(sample_id,Gene,bmi,bacteremia_duration,charleson_index,congestive_heart_failure,
         death_during_admission,history_of_myocardial_infarction,transplant_type,pathogen,icu_admission,
         hypotension,gender,race,smoking,polymicrobial_bacteremia,neutropenia_anc_1500,leukemia_or_lymphoma,Abundance) %>% 
  ungroup()

topn = read_csv("01_r_processed_data/top10_faecalis_v_faecium_proteomics_biomarkers.csv") %>% 
  pull(Gene)

source("functions.R")

stat = metadata_assesment(md_assesment_data,
                          metric = Gene,
                          sample_id = sample_id,
                          features = topn,
                          outcome = Abundance,
                          outcome_c = "Abundance")

  stat_final = rstatix::adjust_pvalue(stat,"p",output.col = "p.adj",method = "fdr")

## Plotting ###
s7a = stat_final %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~Gene,nrow = 2) +
  xlab("metadata_field")

s7a

ggsave("03_supplemental/s7_biomarkers_clinical_metadata_assessment/s7a.pdf",s7a,height = 6, width = 16)
```

### B- Faecalis vs Faecium Metabolomics.

```{r,fig.height = 8, fig.width = 16}
metabolomics_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

m = inner_join(metabolomics_data,gnps_annotations,by = "row_id")

# Defining the input metadata. Need to chose which columns to keep here. 
md_assesment_data= m %>%
  filter(condition != "healthy") %>% 
  select(sample_id,row_id,bmi,bacteremia_duration,charleson_index,congestive_heart_failure,
         death_during_admission,history_of_myocardial_infarction,transplant_type,pathogen,icu_admission,
         hypotension,gender,race,smoking,polymicrobial_bacteremia,neutropenia_anc_1500,leukemia_or_lymphoma,norm_value) %>% 
  ungroup()

topn = read_csv("01_r_processed_data/top10_faecalis_v_faecium_metabolomics_biomarkers.csv") %>% 
  pull(row_id)

source("functions.R")

stat = metadata_assesment(md_assesment_data,
                          metric = row_id,
                          sample_id = sample_id,
                          features = topn,
                          outcome = norm_value,
                          outcome_c = "norm_value")

  stat_final = rstatix::adjust_pvalue(stat,"p",output.col = "p.adj",method = "fdr")

## Plotting ###
s7b = stat_final %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~compound_name_cleaned,nrow = 2) +
  xlab("metadata_field")

s7b

ggsave("03_supplemental/s7_biomarkers_clinical_metadata_assessment//s7b.pdf",s7b,height = 6, width = 16)
```

### C - Proteomic Biomarkers Identified in Live vs Dead.

```{r,fig.height = 8, fig.width = 16}
clinical_md = read_csv("00_r_input_data/clinical_metadata.csv")
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

# Defining the input metadata. Need to chose which columns to keep here. 
md_assesment_data= protein_sum %>% 
  inner_join(clinical_md, by = c("BioReplicate" = "sample_id")) %>% 
  filter(!is.na(death_during_admission)) %>% 
  select(sample_id,Gene,bmi,bacteremia_duration,charleson_index,congestive_heart_failure,
         death_during_admission,history_of_myocardial_infarction,transplant_type,pathogen,icu_admission,
         hypotension,gender,race,smoking,polymicrobial_bacteremia,neutropenia_anc_1500,leukemia_or_lymphoma,Abundance) %>% 
  ungroup()

topn = read_csv("01_r_processed_data/top10_live_v_dead_proteomics_biomarkers.csv") %>% 
  pull(Gene)

source("functions.R")

stat = metadata_assesment(md_assesment_data,
                          metric = Gene,
                          sample_id = sample_id,
                          features = topn,
                          outcome = Abundance,
                          outcome_c = "Abundance")

  stat_final = rstatix::adjust_pvalue(stat,"p",output.col = "p.adj",method = "fdr")

## Plotting ###
s7c = stat_final %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~Gene,nrow = 2) +
  xlab("metadata_field")

s7c

ggsave("03_supplemental/s7_biomarkers_clinical_metadata_assessment//s7c.pdf",s7c,height = 6, width = 16)
```

### D - Metabolomic Biomarker Identified in Live vs Dead

```{r,fig.height = 8, fig.width = 16}
metabolomics_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

m = inner_join(metabolomics_data,gnps_annotations,by = "row_id")

# Defining the input metadata. Need to chose which columns to keep here. 
md_assesment_data= m %>%
  filter(!is.na(death_during_admission)) %>% 
  select(sample_id,row_id,bmi,bacteremia_duration,charleson_index,congestive_heart_failure,
         death_during_admission,history_of_myocardial_infarction,transplant_type,pathogen,icu_admission,
         hypotension,gender,race,smoking,polymicrobial_bacteremia,neutropenia_anc_1500,leukemia_or_lymphoma,norm_value) %>% 
  ungroup()

topn = read_csv("01_r_processed_data/top10_live_v_dead_metabolomics_biomarkers.csv") %>% 
  pull(row_id)

source("functions.R")

stat = metadata_assesment(md_assesment_data,
                          metric = row_id,
                          sample_id = sample_id,
                          features = topn,
                          outcome = norm_value,
                          outcome_c = "norm_value")

  stat_final = rstatix::adjust_pvalue(stat,"p",output.col = "p.adj",method = "fdr")

## Plotting ###
s7d = stat_final %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~compound_name_cleaned,nrow = 2) +
  xlab("metadata_field")

s7d

ggsave("03_supplemental/s7_biomarkers_clinical_metadata_assessment/s7d.pdf",s7d,height = 6, width = 16)

```

## S8 -Cytokine Inference.

Network-based Cytokine Inference Knowledge-based networks were used to infer the relative contributions of major cytokines on the observed proteomic alterations. First, a list of cytokines was manually curated (TGFb, TNF, IFN, IL1-40, CXCL1-16, CCL1-27) and submitted with the significantly altered proteins (uncorrected ANOVA p \< 0.05) to the STRING-db tool (all active interaction sources, interaction score \> 0.4). The network was exported into a simple tabular output and cytokines were profiled for their enrichment in mortality clusters (pMortality+, pMortality and pMortality) using the following approach. First, cytokines were filtered to have at least five connections to mortality networks. Then, the proportion of each cytokine within the mortality clusters (ie. number of cytokine-mortality cluster connections relative to the total connections within mortality clusters) was compared to the proportion of each cytokine in the entire dataset (ie. number of cytokine-protein connections relative to total connections within the entire dataset). Significance was determined using Chi-square tests and a fold-change was calculated by dividing the proportion of cytokine-mortality connections by the proportion of cytokine connections within the total dataset. A final enrichment score was calculated by multiplying the -log10(p value) from the Chisquare test by the log2(fold-change of enrichment). By including the total number of connections that each cytokine had in the entire dataset as a background, any enrichment bias due to higher annotation rates for popular cytokines in STRING-db is controlled for.

### Upstream Data Processing

```{r}
### Loading data ###
faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv")
# Name Mapping 
annotations = read_csv("00_r_input_data/gene_uniprot_annotations.csv") 
#changing the names to uniprotid
faecalis_v_faecium = left_join(faecalis_v_faecium,annotations,by = c("uniprotid"))
# reading in the curated cytokine list
cytokine_list = read_csv("00_r_input_data/jacobs_cytokine_list.csv",skip = 1)

### Cytokine Network generated for all proteins detected in the study ### 
background = faecalis_v_faecium %>% 
  pull(gene_name)

#exporting file that we can use with the stringdb web tool
#https://www.string-db.org/cgi/input?sessionId=bjuiDWoZnTQb&input_page_show_search=on
all_proteins_and_cytokines = c(background,cytokine_list$cytokines)
write_lines(all_proteins_and_cytokines,"01_r_processed_data/cytokine_inference/00_all_proteins_and_cytokines.csv")

# Results were exported from the stringdb web tool. 
# Settings were set as follows: minimum required interaction score = 0.400 (medium),  active interaction sources = all (textmining, neighborhood, experiments, gene fusions, databases, co-occurenece, co-expression)
# Gene names that did not find a map were disregarded. 
# Was exported as tabular text output with reciprocal edges

# Reading in results input into string db using faecalis_out proteins 
string_db_results = read_tsv("01_r_processed_data/cytokine_inference/01_all_proteins_and_cytokines_string_db_out.tsv") %>% 
  #only keeping connections that are between a protein identified and cytokines
  filter(`#node1` %in% background & `node2` %in% cytokine_list$cytokines)

write_csv(string_db_results,"01_r_processed_data/cytokine_inference/02_all_proteins_and_cytokines_string_db_out_filtered.csv")

# counting the number of times a cytokine interacted with a detected protein

n_background_proteins = length(background)

background_cytokines = string_db_results %>% 
  group_by(node2) %>% 
  summarise(n = n()) %>% 
  mutate(condition = "background_present")

write_lines(cytokine_list$cytokines,"01_r_processed_data/cytokine_inference/cytokine_node_list")
```

### A - Overview of entire network

This figure is generated in cytoscape using 02_all_proteins_and_cytokines_string_db_out_filtered.csv as the input network. Biopax style was used. Labels were removed. Background was changed to white and the edges were changed to black with a transparency of 75. Nodes in node 2 were selected using the cytokine_node_list file and their size (40), color (dark red) and the shape (diamond) was changed to distinguish the cytokines from other nodes.

4 cytoscape diagrams were created using the initial plot as a template in order to highlight the protein nodes by the protein belonging to each condition. Keeping with the scheme of the paper, healthy samples were colored yellow, faecalis infected samples were colored navy, and faecium colored samples were colored turquoise.

### B - Explanation of the stats used to identify signficant cytokines

This figure was created in biorender.

### C - Cytokines that were found to be significantly altered among the conditions studied.

#### Infected

```{r,fig.height = 10,fig.width = 5}
## Doing this for Infected vs Background ##

### Now lets try doing this in Infected vs Healthy ### 
infected_vs_healthy = read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
# Name Mapping 
annotations = read_csv("00_r_input_data/gene_uniprot_annotations.csv") 
#changing the names to uniprotid
infected_vs_healthy = left_join(infected_vs_healthy,annotations,by = c("uniprotid"))

# figuring otu what genes were up in infected
up_in_infected = infected_vs_healthy %>% 
  filter(log2FC>0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(gene_name)

n_infected_proteins = length(up_in_infected)

infected_cytokines = string_db_results %>% 
  filter(`#node1` %in% up_in_infected) %>% 
  group_by(node2) %>% 
  summarise(n= n()) %>% 
  mutate(condition = "infected_present")

background_infected = bind_rows(background_cytokines,infected_cytokines) %>% 
  pivot_wider(names_from = condition, values_from = n) %>% 
  mutate(infected_present = replace_na(infected_present,0),
         background_absent = n_background_proteins- background_present,
         infected_absent = n_infected_proteins - infected_present)

# Making a contingency table

out = data.frame()

for(i in unique(background_infected$node2)){

  contingency_table_present = background_infected %>% 
  filter(node2 == i) %>% 
  select(infected_present,background_present) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(present = V1)

  contingency_table_absent =background_infected %>% 
  filter(node2 == i) %>% 
  select(infected_absent,background_absent) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(absent = V1)

  contingency_table = bind_cols(contingency_table_present,contingency_table_absent) %>% 
  as.matrix()

  stat = rstatix::fisher_test(contingency_table) %>% 
    mutate(cytokine = i )

  out = rbind(stat,out)

}

corrected_pvalues = out %>% 
  rstatix::adjust_pvalue(method = "fdr")

# Now calculating fold change

log2fc = background_infected %>% 
  select(node2,
         background_present,
         infected_present) %>% 
  mutate(n_background_proteins = n_background_proteins,
         n_infected_proteins = n_infected_proteins,
         proportion_infected = infected_present/n_infected_proteins,
         proportion_background = background_present/n_background_proteins,
         log2fc = log2(proportion_infected/proportion_background)) %>% 
  select(node2,log2fc,proportion_in_condition = proportion_infected,
         proportion_in_background = proportion_background)

# all metrics
infected_metrics = corrected_pvalues %>% 
  inner_join(log2fc,by = c("cytokine" ="node2")) %>% 
  mutate(enrichment_score = -log10(p.adj) * log2fc) %>% 
  mutate(condition = "infected")

write_lines(up_in_infected,"01_r_processed_data/cytokine_inference/infected_protein_list")

```

#### Healthy

```{r}
### Now lets try doing this in Infected vs Healthy ### 
infected_vs_healthy = read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
# Name Mapping 
annotations = read_csv("00_r_input_data/gene_uniprot_annotations.csv") 
#changing the names to uniprotid
infected_vs_healthy = left_join(infected_vs_healthy,annotations,by = c("uniprotid"))

# up in healthy
up_in_healthy = infected_vs_healthy %>% 
  filter(log2FC<0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(gene_name)

n_healthy_proteins = length(up_in_healthy)

healthy_cytokines = string_db_results %>% 
  filter(`#node1` %in% up_in_healthy) %>% 
  group_by(node2) %>% 
  summarise(n= n()) %>% 
  mutate(condition = "healthy_present")

background_healthy = bind_rows(background_cytokines,healthy_cytokines) %>% 
  pivot_wider(names_from = condition, values_from = n) %>% 
  mutate(healthy_present = replace_na(healthy_present,0),
         background_absent = n_background_proteins- background_present,
         healthy_absent = n_healthy_proteins - healthy_present)

# Making a contingency table

out = data.frame()

for(i in unique(background_healthy$node2)){

  contingency_table_present = background_healthy %>% 
  filter(node2 == i) %>% 
  select(healthy_present,background_present) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(present = V1)

  contingency_table_absent =background_healthy %>% 
  filter(node2 == i) %>% 
  select(healthy_absent,background_absent) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(absent = V1)

  contingency_table = bind_cols(contingency_table_present,contingency_table_absent) %>% 
  as.matrix()

  stat = rstatix::fisher_test(contingency_table) %>% 
    mutate(cytokine = i )

  out = rbind(stat,out)

}

corrected_pvalues = out %>% 
  rstatix::adjust_pvalue(method = "fdr")

# Now calculating fold change

log2fc = background_healthy %>% 
  select(node2,
         background_present,
         healthy_present) %>% 
  mutate(n_background_proteins = n_background_proteins,
         n_healthy_proteins = n_healthy_proteins,
         proportion_healthy = healthy_present/n_healthy_proteins,
         proportion_background = background_present/n_background_proteins,
         log2fc = log2(proportion_healthy/proportion_background)) %>% 
  select(node2,log2fc,proportion_in_condition = proportion_healthy,
         proportion_in_background = proportion_background)

# all metrics
healthy_metrics = corrected_pvalues %>% 
  inner_join(log2fc,by = c("cytokine" ="node2")) %>% 
  mutate(enrichment_score = -log10(p.adj) * log2fc,
         condition = "healthy")

write_lines(up_in_healthy,"01_r_processed_data/cytokine_inference/healthy_protein_list")

```

#### Faecalis

```{r}
### Now lets try doing this in Facalis vs Faecium ### 
faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv")

# Name Mapping 
annotations = read_csv("00_r_input_data/gene_uniprot_annotations.csv") 
#changing the names to uniprotid
faecalis_v_faecium = left_join(faecalis_v_faecium,annotations,by = c("uniprotid"))

# up in faecalis
up_in_faecalis = faecalis_v_faecium %>% 
  filter(log2FC>0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(gene_name)

n_faecalis_proteins = length(up_in_faecalis)

faecalis_cytokines = string_db_results %>% 
  filter(`#node1` %in% up_in_faecalis) %>% 
  group_by(node2) %>% 
  summarise(n= n()) %>% 
  mutate(condition = "faecalis_present")

background_faecalis = bind_rows(background_cytokines,faecalis_cytokines) %>% 
  pivot_wider(names_from = condition, values_from = n) %>% 
  mutate(faecalis_present = replace_na(faecalis_present,0),
         background_absent = n_background_proteins- background_present,
         faecalis_absent = n_faecalis_proteins - faecalis_present)

# Making a contingency table

out = data.frame()

for(i in unique(background_faecalis$node2)){

  contingency_table_present = background_faecalis %>% 
  filter(node2 == i) %>% 
  select(faecalis_present,background_present) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(present = V1)

  contingency_table_absent = background_faecalis %>% 
  filter(node2 == i) %>% 
  select(faecalis_absent,background_absent) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(absent = V1)

  contingency_table = bind_cols(contingency_table_present,contingency_table_absent) %>% 
  as.matrix()

  stat = rstatix::fisher_test(contingency_table) %>% 
    mutate(cytokine = i )

  out = rbind(stat,out)

}

corrected_pvalues = out %>% 
  rstatix::adjust_pvalue(method = "fdr")

# Now calculating fold change

log2fc = background_faecalis %>% 
  select(node2,
         background_present,
         faecalis_present) %>% 
  mutate(n_background_proteins = n_background_proteins,
         n_faecalis_proteins = n_faecalis_proteins,
         proportion_faecalis = faecalis_present/n_faecalis_proteins,
         proportion_background = background_present/n_background_proteins,
         log2fc = log2(proportion_faecalis/proportion_background)) %>% 
  select(node2,log2fc,proportion_in_condition = proportion_faecalis,
         proportion_in_background = proportion_background)

# all metrics
faecalis_metrics = corrected_pvalues %>% 
  inner_join(log2fc,by = c("cytokine" ="node2")) %>% 
  mutate(enrichment_score = -log10(p.adj) * log2fc,
         condition = "faecalis")

write_lines(up_in_faecalis,"01_r_processed_data/cytokine_inference/faecalis_protein_list")

```

#### Faecium

```{r}
### Now lets try doing this in Facalis vs Faecium ### 
faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv")

# Name Mapping 
annotations = read_csv("00_r_input_data/gene_uniprot_annotations.csv") 
#changing the names to uniprotid
faecalis_v_faecium = left_join(faecalis_v_faecium,annotations,by = c("uniprotid"))

# up in faceium 
up_in_faecium = faecalis_v_faecium %>% 
  filter(log2FC<0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  arrange(adj.pvalue) %>% 
  pull(gene_name)

n_faecium_proteins = length(up_in_faecium)

faecium_cytokines = string_db_results %>% 
  filter(`#node1` %in% up_in_faecium) %>% 
  group_by(node2) %>% 
  summarise(n= n()) %>% 
  mutate(condition = "faecium_present")

background_faecium = bind_rows(background_cytokines,faecium_cytokines) %>% 
  pivot_wider(names_from = condition, values_from = n) %>% 
  mutate(faecium_present = replace_na(faecium_present,0),
         background_absent = n_background_proteins- background_present,
         faecium_absent = n_faecium_proteins - faecium_present)

# Making a contingency table

out = data.frame()

for(i in unique(background_faecalis$node2)){

  contingency_table_present = background_faecium %>% 
  filter(node2 == i) %>% 
  select(faecium_present,background_present) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(present = V1)

  contingency_table_absent = background_faecium %>% 
  filter(node2 == i) %>% 
  select(faecium_absent,background_absent) %>% 
  t() %>% 
  as.data.frame() %>% 
  rename(absent = V1)

  contingency_table = bind_cols(contingency_table_present,contingency_table_absent) %>% 
  as.matrix()

  stat = rstatix::fisher_test(contingency_table) %>% 
    mutate(cytokine = i )

  out = rbind(stat,out)

}

corrected_pvalues = out %>% 
  rstatix::adjust_pvalue(method = "fdr")

# Now calculating fold change

log2fc = background_faecium %>% 
  select(node2,
         background_present,
         faecium_present) %>% 
  mutate(n_background_proteins = n_background_proteins,
         n_faecium_proteins = n_faecium_proteins,
         proportion_faecium= faecium_present/n_faecium_proteins,
         proportion_background = background_present/n_background_proteins,
         log2fc = log2(proportion_faecium/proportion_background)) %>% 
  select(node2,log2fc,proportion_in_condition = proportion_faecium,
         proportion_in_background = proportion_background)

# all metrics
faecium_metrics = corrected_pvalues %>% 
  inner_join(log2fc,by = c("cytokine" ="node2")) %>% 
  mutate(enrichment_score = -log10(p.adj) * log2fc,
         condition = "faecium")

write_lines(up_in_faecium,"01_r_processed_data/cytokine_inference/faecium_protein_list")
```

### D - Plot of stats

```{r,fig.height = 12, fig.width = 5}
all_metrics = dplyr::bind_rows(infected_metrics,healthy_metrics,faecalis_metrics,faecium_metrics) %>% 
  mutate(direction_of_change = case_when(log2fc > 0 ~ "increased",
                                         TRUE ~ "decreased"),
         )

all_metrics$condition = factor(all_metrics$condition, levels=c("infected","healthy","faecalis","faecium"))

s8d = all_metrics %>% 
  ggplot(aes(reorder(cytokine, -p.adj),p.adj,color = condition,size = direction_of_change))+
  geom_point()+
  coord_flip()+
  geom_hline(yintercept = 0.05,linetype = "dashed",color = "red")+
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 90))+
  xlab("Cytokine")+
  scale_colour_manual(values = c("#FFA500","#FDE725FF","#440154FF","#2A788EFF"))

s8d

ggsave("03_supplemental/s8_cytokine_inference/s8d.pdf",plot = s8d,width = 5,height = 12,units = "in")
```

### E - Network of infected

Figure plotted via cytoscape.

```{r}
# Since we have 
cytokines_in_infected_network = infected_metrics %>% 
  filter(p < 0.05) %>% 
  pull(cytokine)

#writing a file so we can look at the network further. 
infected_cytokine_network = string_db_results %>% 
  filter(node2 %in% cytokines_in_infected_network,
         `#node1` %in% up_in_infected)

write_csv(infected_cytokine_network,"01_r_processed_data/cytokine_inference/03_infected_protein_cytokine_network_filtered.csv")
```

## S9 - Prediction of Clinical Outcome by Nanopore Sequencing

### A - Sequencing Data from Strains of faecalis, colored by clinical outcome

```{r,fig.height = 8, fig.width = 12}
## Loading in all the data ### 
md = read_csv("00_r_input_data/clinical_metadata.csv")
gene_data = read_csv("00_r_input_data/gene_presence_absence_long.csv")

## Combining data and metadata ## 
gene_combined = inner_join(md,gene_data,by = "sample_id")

# Formating ##
faecalis = gene_combined %>% 
  filter(condition == "faecalis") %>% 
  filter(!is.na(death_during_admission)) %>% 
  select(sample_id,death_during_admission,Gene,value) %>% 
  pivot_wider(names_from = Gene,values_from = value)

## Getting into a format where we can do the clustering ## 
faecalis_data = faecalis[,3:28290] %>% 
   #Remove columns that only contain 0s 
  select_if(colSums(.) != 0) %>% 
  as.data.frame()

## Fixing the row names so they match for the heatmap ##
rownames(faecalis_data) = faecalis$sample_id

faecalis_annotation = faecalis[,2] %>% 
  as.data.frame()

rownames(faecalis_annotation) = faecalis$sample_id

## Generating a dendrogram without a heatmap ##
library(dendextend)

d = dist(faecalis_data,method = "binary")

dend = as.dendrogram(hclust(d, method = "ward.D2"))

# let's add some color: Red = dead, black = okay
colors_to_use = c("#000000","#FF0000")[as.numeric(as.factor(faecalis_annotation$death_during_admission))]

#getting in the same order as dendrogram
colors_to_use = colors_to_use[order.dendrogram(dend)]

# Now we can color the brancehes
dend = branches_color(dend,col = colors_to_use,) %>% 
  set("branches_lwd", 8)

library(pheatmap)
## Generating a dendogram with a heatmap ##
faecalis_annotation = faecalis_annotation %>% 
  mutate(death_during_admission = as.character(death_during_admission)) %>% 
  select(mortality = death_during_admission)

#plotting with a heatmap
p = pheatmap(faecalis_data,
         clustering_distance_rows = "binary",
         clustering_method = "ward.D2",
         annotation_row = faecalis_annotation,
         labels_col=rep("",ncol(faecalis_data)))

save_pheatmap_pdf <- function(x, filename, width=12, height=8) {
  pdf(filename, width = width, height = height)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}

save_pheatmap_png <- function(x, filename, width=12, height=8,res = 150) {
  png(filename, width = width, height = height,res = 150,units = "in")
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}

#Have trouble manipulating the pdf in inkscape
save_pheatmap_pdf(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.pdf")
save_pheatmap_png(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.png")

```

### B - Sequencing Data from Faecium strains, colored by clinical outcome

```{r,fig.height = 8, fig.width = 12}
## Processing the data to only contain the faecium isolates ###
faecium = gene_combined %>% 
  filter(condition == "faecium") %>% 
  filter(!is.na(death_during_admission)) %>% 
  select(sample_id,death_during_admission,Gene,value) %>% 
  pivot_wider(names_from = Gene,values_from = value) %>% 
  # Removing weird sample that was found to be E.coli.
  filter(sample_id != "S68")

# Processing to get rid of columns we dont need
faecium_data = faecium[,3:28290] %>% 
   #Remove columns that only contain 0s 
  select_if(colSums(.) != 0) %>% 
  as.data.frame()

# Changing the row names so we can plot
rownames(faecium_data) = faecium$sample_id

# Extracting the annotations 
faecium_annotation = faecium[,2] %>% 
  as.data.frame() %>% 
  mutate(death_during_admission = as.character(death_during_admission)) %>% 
  select(mortality = death_during_admission)

rownames(faecium_annotation) = faecium$sample_id

p2 = pheatmap(faecium_data,
         clustering_distance_rows = "binary",
         clustering_method = "ward.D2",
         annotation_row = faecium_annotation,
         labels_col=rep("",ncol(faecium_data)))

save_pheatmap_pdf(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.pdf")
save_pheatmap_png(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.png")
```

## S10 - Checking Confounding Variables

### Faecalis vs Faecium

#### A- IG abundances between patients with no transplant

```{r}
md = read_csv("00_r_input_data/clinical_metadata.csv")

protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

anot = read_csv("00_r_input_data/all_protein_annotations.csv")

sum_annotated = left_join(protein_sum,anot,by = c("uniprotid" = "Entry")) %>% 
  left_join(md,by = c("BioReplicate" = "sample_id")) %>% 
  filter(transplant_type == "none") %>% 
#Removing one problematic immunglobulin. Negative values for 3 of the samples messed things up. 
  #Doesn't change the story, just the plot. 
  filter(uniprotid != "A0A0B4J2B5")

ig = sum_annotated %>% 
  filter(grepl(".*Immunoglobulin.*",Protein.names)) %>% 
  filter(!grepl(".*receptor.*",Protein.names))

stat = ig %>% 
  rstatix::t_test(Abundance ~ Condition) %>% 
  rstatix::add_significance()

s10a = ig %>% 
  ggplot(aes(Condition,Abundance))+
  geom_violin()+
  geom_boxplot()+
  ggpubr::stat_pvalue_manual(stat,
                             y.position = 22,
                             step.increase = 0.1)+
  theme_classic()

s10a

ggsave("03_supplemental/s10_checking_confounding_variables/s10a.pdf",s10a)
```

#### B- Between Smoking and Non Smoking

```{r}

smoking_associated = c("APOC3","IGHV3-7","IGLV1-47","RBP4")
md = read_csv("00_r_input_data/clinical_metadata.csv")

protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

anot = read_csv("00_r_input_data/all_protein_annotations.csv")

sum_annotated = left_join(protein_sum,anot,by = c("uniprotid" = "Entry")) %>% 
  left_join(md,by = c("BioReplicate" = "sample_id")) %>% 
  filter(Gene %in% smoking_associated,
         condition != "healthy",
         smoking == FALSE)

plotlist = list()

for(i in smoking_associated){

  f = filter(sum_annotated,Gene %in% i)

stat = f %>% 
  rstatix::t_test(Abundance ~ condition) %>% 
  rstatix::add_significance() %>% 
  rstatix::add_xy_position()

p1 = ggplot(f,aes(condition,Abundance)) + 
  geom_violin()+
  geom_point()+
  ggtitle(i)+
  ggpubr::stat_pvalue_manual(stat)

plotlist[[i]] = p1

}

library(gridExtra)

s10b = do.call("grid.arrange", plotlist)

ggsave("03_supplemental/s10_checking_confounding_variables/s10b.pdf",s10b)
```

# Getting exact numbers for manuscript writing

## How many of each patient type do we have?

```{r}
md = readr::read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  group_by(condition) %>% 
  summarise(n = n())

md

md2 = readr::read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  group_by(death_during_admission) %>% 
  summarise(n = n())

md2
```

## How many proteins did we identify?

```{r}
p = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  group_by(uniprotid) %>% 
  summarise(n = n())
```

## How many proteins did we identify in all samples

```{r}
p = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  group_by(uniprotid) %>% 
  summarise(n = n())

sum(p$n == 105)
```

## How many metabolites were differentially abundant between healthy and infected?

```{r}
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv",na = c("N/A","NA"))
## Reading in the normalized metabolomics data ##
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>%
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ infection_status) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Extracting the values for each feature for healthy patients
healthy = normalized_data %>% 
  filter(infection_status == "uninfected") %>% 
  group_by(row_id) %>% 
  summarise(healthy = mean(norm_value))

# Extracting the values for each feature for infected patients
infected = normalized_data %>% 
  filter(infection_status == "infected") %>% 
  group_by(row_id) %>% 
  summarise(infected = mean(norm_value))

## Combining the healthy and infected so we can caclulated the log2fc ##
log2fc = inner_join(healthy,infected,by = "row_id") %>% 
  mutate(log2fc_infected_healthy = log2(infected/healthy))

## Joining the statistics and log2 fold change ##  
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE),
         direction = case_when(log2fc_infected_healthy > 0 ~ "up",
                   TRUE ~ "down"))

sum = vp %>% 
  filter(p.adj_fdr < 0.05) %>% 
  group_by(direction) %>% 
  summarise(n = n())

sum
```

## How many proteins

```{r}
faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  mutate(direction = case_when(log2FC > 0 ~ "up",
                   TRUE ~ "down")) %>% 
  filter(adj.pvalue < 0.05) %>% 
  group_by(direction) %>% 
  summarise(n = n())

faecalis_v_faecium
```

## How unbalanced are the faecium vs faecalis group in terms of metadata?

```{r}
md = read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  filter(pathogen != "healthy")

sum = md %>% 
  group_by(pathogen,smoking) %>% 
  summarise(n = n())

sum

sum2 = md %>% 
  group_by(pathogen,transplant_type) %>% 
  summarise(n = n())

sum
```

## Number of proteins different betweeen metabolomics - faecalis faecium

```{r}

## Reading in Data
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>%
    filter(condition %in% c("faecalis","faecium")) %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ condition) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,condition,norm_value)

# Extracting just the faecalis data
faecalis = log2fc_data %>% 
  filter(condition == "faecalis") %>% 
  group_by(row_id) %>% 
  summarise(faecalis = mean(norm_value))

# Extracting just the faecium data 
faecium = log2fc_data %>% 
  filter(condition == "faecium") %>% 
  group_by(row_id) %>% 
  summarise(faecium = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(faecalis,faecium,by = "row_id") %>% 
  mutate(log2fc_faecalis_faecium = log2(faecalis/faecium))

# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE),
         direction = case_when(log2fc_faecalis_faecium > 0 ~ "positive",
           TRUE ~ "negative"
         ))

sum = vp %>% 
  filter(p.adj_fdr <= 0.05) %>% 
  group_by(direction) %>% 
  summarise(n = n())

sum
```

## Number of proteins differentally abundant in mortality vs survival

```{r}
sum = read_csv("01_r_processed_data/death_vs_life.csv") %>% 
  filter(is.na(issue)) %>% 
  mutate(direction = case_when(log2FC > 0 ~ "up",
                               TRUE ~ "down")) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  group_by(direction) %>% 
  summarise(n = n())

sum
```

## Number of metabolites differentially abundant in mortality vs survival

```{r}
## Reading in Data
normalized_data = read_csv("01_r_processed_data/normalized_metabolomics.csv") %>% 
  filter(!is.na(death_during_admission))

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ death_during_admission) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,death_during_admission,norm_value)

# Extracting just the faecalis data
dead = log2fc_data %>% 
  filter(death_during_admission == TRUE) %>% 
  group_by(row_id) %>% 
  summarise(dead = mean(norm_value))

# Extracting just the faecium data 
live = log2fc_data %>% 
  filter(death_during_admission == FALSE) %>% 
  group_by(row_id) %>% 
  summarise(live = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(dead,live,by = "row_id") %>% 
  mutate(log2fc_live_dead = log2(live/dead))

# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE)) %>% 
    mutate(direction = case_when(log2fc_live_dead > 0 ~ "up",
                               TRUE ~ "down")) %>% 
  filter(p.adj_fdr <= 0.05)

sum = vp %>% 
  group_by(direction) %>% 
  summarise(n = n())

sum

```

# Stuff I might end up needing

## PTM analysis of proteomics data via open search.

```{r,fig.height = 10, fig.width = 10}
open_search_results = read_tsv("00_r_input_data/global.modsummary.tsv") %>% 
  select(Modification, `Mass Shift`,!contains("percent")) %>% 
  mutate(all_plexes_sum = (plex1_PSMs+plex2_PSMs+plex3_PSMs+plex4_PSMs+plex5_PSMs+plex6_PSMs+plex7_PSMs),
    all_PSMs = sum(plex1_PSMs,plex2_PSMs,plex3_PSMs,plex4_PSMs,plex5_PSMs,plex6_PSMs,plex7_PSMs),
    all_plexes_percent = all_plexes_sum/all_PSMs * 100,
    mass_modification = paste0(`Mass Shift`,"-",Modification))

open_1_percent = open_search_results %>% 
  filter(all_plexes_percent >= 1) %>% 
  #TMT pro
  mutate(Modification = case_when(Modification == 	
"Representative mass and accurate mass for 113, 114, 116 & 117" ~ "TMTpro",
TRUE ~ Modification),
Modification = case_when(!grepl(".*[1-9].*",Modification) ~ paste0(Modification,"-",round(`Mass Shift`, 2)),
                         TRUE ~ Modification))

p1 = open_1_percent %>% 
  ggplot(aes(reorder(Modification,all_plexes_percent),all_plexes_percent))+
  geom_col()+
  coord_flip()+
  ylab("Percent PSMs")+
  xlab("Putative Identification")

p1  
```

## Combined Proteomics and Metabolomics clustering of mortality.

```{r,fig.height = 10, fig.width = 12,eval = FALSE}
##Proteomics data ##
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  select(tmt_id = sample_id,everything())

## Clinical metadata ##
map = readxl::read_excel("00_r_input_data/proteomics_sample_mapping.xlsx") %>% 
  mutate(tmt_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(tmt_id,sample_id = sample_name)

md = read_csv("00_r_input_data/clinical_metadata.csv")

md_final = inner_join(map,md,by = "sample_id")

clean_clinical = inner_join(pl,md_final, by = "tmt_id")

clean_clinical_wide = clean_clinical %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #filter(sample_id.y != "S49",
         #sample_id.y != "S76") %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame()

## Metabolomics data
## Reading in the normalized data ## 
normalized_data = read_csv("01_r_processed_data//normalized_metabolomics.csv")
gnps_annotations = read_csv("00_r_input_data/gnps_annotations.csv")

## Switching to the wide format ### 
norm_wide = normalized_data %>%
  select(-row_m_z,-row_retention_time,-value) %>% 
  tidyr::pivot_wider(names_from = row_id,values_from = norm_value)

## Looking for only columns that were annotated ## 
gnps_and_in_norm = intersect(unique(gnps_annotations$row_id), colnames(norm_wide))

gnps_annotated_metabolites = norm_wide %>% 
  as.data.frame() %>% 
  select(sample_id,death_during_admission,gnps_and_in_norm) %>% 
  filter(!is.na(death_during_admission))

### Combining all data ### 
all= inner_join(clean_clinical_wide,gnps_annotated_metabolites,by = "sample_id")

# Getting into a format where we can do the clustering ## 
clust_data = all[,2:491] %>% 
   #Remove columns that only contain 0s 
  select_if(colSums(.) != 0) %>% 
  as.data.frame()

## Fixing the row names so they match for the heatmap ##
rownames(clust_data) = clust_data$sample_id

annotation = all[,2] %>% 
  as.data.frame()

## Generating a dendrogram without a heatmap ##
library(dendextend)

#scaling so that proteomics and metabolomics are in terms of same units
scaled = scale(clust_data)
d = dist(scaled)

dend = as.dendrogram(hclust(d, method = "ward.D2"))

# let's add some color: Red = dead, black = okay
colors_to_use = c("#000000","#FF0000")[as.numeric(as.factor(all$death_during_admission))]

#getting in the same order as dendrogram
colors_to_use = colors_to_use[order.dendrogram(dend)]

# Now we can color the brancehes
dend = branches_color(dend,col = colors_to_use,) %>% 
  set("branches_lwd", 8)

plot(dend)

dend %>% rect.dendrogram(k=2, 
                           border = 8, lty = 5, lwd = 2)

xtab <- as.table(rbind(c(4, 8), c(13, 53)))
dimnames(xtab) <- list(
  group = c("k1", "k2"),
  death = c("yes", "no")
)

fisher_test(xtab)
```

## Combined Proteomic and Metabolomic Model to predict Faecalis vs Faecium Bacteremia.

## A - Description of model design.

This figure was generated in Biorender.

## B - Combined Model Performance.

<https://www.tidymodels.org/start/case-study/>

```{r,eval = FALSE}
## Loading in all the data needed for the model ## 
mapping = readxl::read_excel("../proteomics/sample_mapping.xlsx") %>% 
  mutate(sample_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(sample_id,sample_name)

proteomics = read_csv("../Analysis/quant_msstats_protein_level_data.csv") %>% 
  separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>% 
  inner_join(protein_annotations,by = c("uniprotid" = "Entry")) %>% 
  mutate(sample_id = paste0(Mixture,".",Channel)) %>% 
  inner_join(mapping,by = "sample_id") %>% 
  select(sample_name,uniprotid,Abundance) %>% 
  pivot_wider(names_from = uniprotid,values_from = Abundance)

gnps_annotations = read_csv("gnps_annotations.csv")

metabolomics = read_csv("../Manuscript/normalized_metabolomics.csv") %>% 
  select(sample_name = sample_id,condition,row_id,norm_value) %>% 
  # Only looking at the metabolites that we were able to identify
  #filter(row_id %in% gnps_annotations$row_id) %>% 
  pivot_wider(names_from = row_id,values_from = norm_value)

all = inner_join(proteomics,metabolomics,by= "sample_name") %>% 
  select(sample_name,condition,everything()) %>% 
  filter(condition %in% c("faecalis","faecium"))

write_csv(all, "all_features_for_lasso_regression.csv")

```

### All predictors

```{r,eval=FALSE}
### Reading in data for the regression ### 
all = read_csv("all_features_for_lasso_regression.csv")

## What values have less than 50% of the features?  
missing_count = colSums(is.na(all)/nrow(all) * 100) 
missing_g_50 = names(missing_count[missing_count>50])

## Removing features missing > 50% of values ## 
all_50_missing_removed = all %>% 
  select(!contains(missing_g_50))

# Setting seed to make steps relying on randomness reproducible
set.seed(123)

## Imputing values using missranger package ##
library(missRanger)
all_na_impute = missRanger(all_50_missing_removed)

### Performing Lasso Regression ### 
library(tidymodels)

# Splitting the original dataset
split = initial_split(all_na_impute, strata = condition)

# Putting all data in the training set. Did this because we don't have enough samples for classic training/ validation set. 
train = training(split)

test = testing(split)

## Defining the Cross Validation Folds ##
folds = vfold_cv(train, v = 10)

## Defining a recipe. ##
recipe = recipe(condition ~ ., data = train) %>%
  # Removing the ID column from consideration
  update_role(sample_name, new_role = "ID") %>%
  # Getting rid of all variables with zero variance (if they exist)
  step_zv(all_numeric(), -all_outcomes()) %>%
  # Normalizing data to work with the lasso regression. 
  step_normalize(all_numeric(), -all_outcomes())

# making sure strings are strings and not factors. 
prep = recipe %>%
  prep(strings_as_factors = FALSE)

# Defining the lasso model. Here the penalty is a tuned feature. 
# Mixture specifies we should use L1 regularization, which equates to lasso regression. 
lasso_spec = logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

# Making the recipe into a workflow. 
wf = workflow() %>%
  # Processes the data
  add_recipe(recipe) %>% 
  # Adds the model
  add_model(lasso_spec)

## Creating tuning grid ## 
# 0.001 to 0.3 by steps of 0.01, as in Jacob's paper
lr_reg_grid = tibble(penalty = seq(from = 0.001, to = 0.3, by = 0.01))

## Tuning the model to find best penalty to use ##
lr_res = wf %>% 
  tune_grid(resample = folds,
            grid = lr_reg_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

## Looking at how the penalty affects the performance of the model ##
lr_plot = lr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number())

lr_plot

# Extracting the tuning parameter that worked the best #
lr_best = lr_res %>% 
  select_best("roc_auc")

## Updating workflow to contain results from tuning ## 
lr_auc = lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(condition, .pred_faecalis) %>% 
  mutate(model = "multi-omics")

autoplot(lr_auc)

##Doing this with test set ## 
final_wf = wf %>% 
  finalize_workflow(lr_best)

final_fit <- 
  final_wf %>%
  last_fit(split)

# Evaluating on the test set 
final_fit %>%
  collect_predictions() %>% 
  roc_curve(condition, .pred_faecalis) %>% 
  autoplot()
```

### Just metabolomics

```{r,eval = FALSE}
### Performing Lasso Regression ### 
library(tidymodels)

metabolomics_all_na_impute = all_na_impute %>% 
  select(starts_with("RID"))

set.seed(2)
# Splitting the original dataset
split = initial_split(all_na_impute, strata = condition)
#Putting all data as training for now
train = training(split,prop = 1)

# Cross validation folds
folds = vfold_cv(train, v = 10)

# Defining a recipe. 
recipe = recipe(condition ~ ., data = train) %>%
  update_role(sample_name, new_role = "ID") %>%
  step_zv(all_numeric(), -all_outcomes()) %>%
  step_normalize(all_numeric(), -all_outcomes())

prep = recipe %>%
  prep(strings_as_factors = FALSE)

# Defining the lasso model. 
lasso_spec = logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

# Making the recipe into a workflow. 
wf = workflow() %>%
  add_recipe(recipe) %>% 
  add_model(lasso_spec)

# Training the model
lasso_fit = wf %>%
  fit(data = train)

## Creating tuning grid ## 
# 0.001 to 0.3 by steps of 0.01, as in Jacob's paper
lr_reg_grid = tibble(penalty = seq(from = 0.001, to = 0.3, by = 0.01))

## Getting the tuning result ##
lr_res = wf %>% 
  tune_grid(folds,
            grid = lr_reg_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))

## Looking at how the penalty affects the performance of the model  ##
lr_plot = lr_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number())

lr_plot

#Extracting the tuning parameter that worked the best
lr_best = lr_res %>% 
  show_best("roc_auc", n = 1)

## Looking at the ROC Plot using lasso regression with all the predictors in the study ##
lr_auc = lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(condition, .pred_faecalis) %>% 
  mutate(model = "Logistic Regression")

autoplot(lr_auc)

```

## Combining all Models

```{r,eval = FALSE}
# Bringing in the model from everything 
rf_auc = rf_res %>% 
  collect_predictions(parameters = rf_best) %>% 
  roc_curve(children, .pred_children) %>% 
  mutate(model = "Random Forest")

# Bringing in the model from just metabolomics

# Bringing in the model from just proteomics

# Combining and plotting all models.

all_models = bind_rows(rf_auc, lr_auc) %>% 
  ggplot(aes(x = 1 - specificity, y = sensitivity, col = model)) + 
  geom_path(lwd = 1.5, alpha = 0.8) +
  geom_abline(lty = 3) + 
  coord_equal() + 
  scale_color_viridis_d(option = "plasma", end = .6)
```

## How many features did we detect as a whole?

```{r,eval = FALSE}
proteomics_count = read_tsv("../proteomics/fragpipe_output_fixed_labels/tmt-report/abundance_gene_MD.tsv") %>% 
  nrow()

conserved_all_samples_proteomics_count = read_tsv("../proteomics/fragpipe_output_fixed_labels/tmt-report/abundance_gene_MD.tsv") %>% 
  select(6:length(.)) %>% 
  na.omit() %>% 
  nrow()

metabolite_features_count = read_csv("../metabolomics/mzmine_output/features_quant.csv") %>% 
  nrow()

## finding out how many annotated features there are. 
gnps_annotations = read_csv("gnps_annotations.csv")

gnps_annotations_n = nrow(gnps_annotations)

## How many of these annotated features were found in every sample?

gnps_annotations_all = read_csv("unnormalized_metabolomics.csv") %>% 
  filter(row_id %in% gnps_annotations$row_id) %>% 
  filter(value > 0 ) %>% 
  group_by(row_id) %>% 
  summarise(n = n()) %>% 
  filter(n == 105) %>% 
  nrow()
```

```{r,eval = FALSE}
#reading in the data
pl = read_csv("../Analysis/quant_msstats_protein_level_data.csv")

clean = pl %>% 
  tidyr::separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>% 
  dplyr::group_by(Mixture,uniprotid,Channel,Condition) %>% 
  dplyr::summarise(sum = sum(Abundance))

infected_healthy_l = clean %>% 
  mutate(classlabels = case_when(Condition %in% c("Faecalis","Faecium") ~ 1,
                                   TRUE ~ 0))

infected_healthy_w = infected_healthy_l %>% 
  pivot_wider(names_from = uniprotid,
              values_from = sum) %>% 
  mutate(sampleid = paste0(Mixture,".",Channel)) %>% 
  ungroup() %>% 
  dplyr::select(-Mixture,-Channel,-Condition,-sampleid) %>% 
  select_if(~ !any(is.na(.))) %>% 
  as.data.frame()

#need to come up with some sort of way to combine all of the fractions and plexes
# The problem is that na.omit filters everything out. 
set.seed(2)

ensemble_result = EFS::ensemble_fs(infected_healthy_w,1,cor_threshold = 0)

long = ensemble_result %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("model") %>% 
  pivot_longer(2:length(.))

top = long %>% 
  group_by(name) %>% 
  dplyr::summarise(sum = sum(value)) %>% 
  dplyr::arrange(-sum) %>% 
  mutate(rank = dense_rank(desc(sum))) %>% 
  select(rank,everything()) %>% 
  filter(name %in% known_indicatiors_of_bacterial_infection)

```

```{r,eval = FALSE}
protein_annotations = read_csv("../Analysis/all_protein_annotations.csv")

protein_sum = read_csv("../Analysis/quant_msstats_protein_level_data.csv") %>% 
   separate(Protein, c("orientation","uniprotid","locus"),sep ="\\|") %>% 
   inner_join(protein_annotations,by = c("uniprotid" = "Entry")) %>% 
  mutate(sample_id = paste0(Mixture,".",Channel))

topn = top %>% 
  pull(name)

ROC_data = dplyr::filter(protein_sum,uniprotid %in% topn) %>% 
  dplyr::mutate(Condition = case_when(Condition %in% c("Faecalis","Faecium") ~ "infected",
                                      TRUE ~ Condition))

make_roc = function(ROC_data){

  list = list()

    for(i in topn){

      data = filter(ROC_data, uniprotid == i)

        mod = glm(as.factor(Condition) ~ Abundance, 
            data = data,
            family = "binomial")

      predicted <- predict(mod ,family = "binomial", data, type="response")

      #define object to plot
      rocobj <- pROC::roc(data$Condition, predicted)

      auc = round(as.numeric(gsub(".*: ", "",rocobj$auc)),2)
#create ROC plot
p1 = pROC::ggroc(rocobj)+
  ggtitle(paste0(i," AUC: ",auc))

list[[i]] = p1
    }
  return(list)
  }

l = make_roc(ROC_data)

f2e = do.call(ggarrange,l)

f2e
```

## Average missingness for lasso regression

Note that we got rid of all the missing data for the metabolomics data as part of the normalization used.

```{r,eval = FALSE}
# All combined. 
all = read_csv("all_features_for_lasso_regression.csv") %>% 
  select(3:length(.))

mean(colSums(is.na(all))/nrow(all) * 100)

# Looking at just metabolomics.

metabolomics = all %>% 
  select(dplyr::starts_with("RID"))

# Sum of na/ total rows * 100 = percent NA.
mean(colSums(is.na(metabolomics))/nrow(metabolomics) * 100)

# Just proteomics.

proteomics = all %>% 
  select(!dplyr::starts_with("RID"))

# Sum of na/ total rows * 100 = percent NA.
mean(colSums(is.na(proteomics))/nrow(proteomics) * 100)

```

## Annotating Proteins

```{r,eval = FALSE}
annotations = read_delim("../proteomics/fragpipe_output_fixed_labels/tmt-report/abundance_gene_MD.tsv") %>% 
  select(gene_name = Index,uniprotid = ProteinID)

write_csv(annotations,"gene_uniprot_annotations.csv")

```

PTM Analysis - GNPS

```{r,fig.height = 10, fig.width = 30}
## Reading in the PTM data from GNPS ## 
ptm_data = read_delim("../metabolomics/ProteoSAFe-METABOLOMICS-SNETS-V2-3e73be4d-view_edges_delta_histogram/networkedges_selfloop/ae6d9d27f8684ad39b8e73b2d6f3da2e..selfloop",na = " ") %>% 
  mutate(DeltaMZ = abs(DeltaMZ))

  # filtering to only include the data in group
  data_f = ptm_data %>% 
    filter(#DeltaMZ >= min & DeltaMZ <= max,
           DeltaMZ > 0)

  # Extracting the annotations from ptm data 
  annotations = data_f %>% 
    group_by(EdgeAnnotation) %>% 
    summarise(mean = mean(DeltaMZ), n = n()) %>% 
    filter(n >= 10,
           !is.na(EdgeAnnotation))

  max_annotation_value = annotations %>% 
    pull(mean) %>% 
    max()

  # plotting 
  p1 = data_f %>% 
    filter(DeltaMZ <= max_annotation_value + 20) %>% 
    #filter(group == group) %>% 
     ggplot(aes(DeltaMZ))+
    geom_histogram(binwidth = 1)+
    ggrepel::geom_text_repel(data = annotations, aes(x = mean, y = 1000,label = EdgeAnnotation, angle = 90),max.overlaps = 1000)

  p1

```

GNPS table

```{r}
arranged_annotations = dplyr::arrange(annotations,-n)

DT::datatable(arranged_annotations)

#Making a plot of the table
pdf("annotations.pdf",height = 4, width = 4)
gridExtra::grid.table(arranged_annotations)
dev.off()
```

```{r}
#total number of MS2 scans
ms2_scans = read_delim("../proteomics/gnps/clusterinfo/93fdf96f02e44389b503c53353712f6a.clusterinfo") %>% 
  group_by(`#Filename`) %>% 
  summarise(n = n())

27017 + 16665
```