### 2024-10-04 max_analysis/model_workflow.qmd
Added cleaned files to my directory. Also added .qmd with model workflow

---
title: "Clinical Data Model"
format: html
editor: visual
---

```{r}
#| label: load-libraries
#| echo: false
#| include: false
library(dplyr)
library(tidymodels)
library(caret)
library(broom)
```

```{r}
#| label: load-clinical-metadata
#| echo: false
#| include: false
clinical_metadata <- read.csv('clinical_metadata.csv')
head(clinical_metadata)
```

```{r}
#| label: filter-clinical
#| echo: false
#| include: true
clinical_edited <- clinical_metadata %>%
  select(sample_id, patient_id, condition, infection_status)
clinical_edited
```

```{r}
#| label: load-proteomics
#| echo: false
#| include: true
proteomics_data <- read.csv('proteomics_data_cleaned.csv')

proteomics_data
```

```{r}
#| label: transpose-proteomics
#| echo: false
#| include: true
proteomics_t <- as.data.frame(t(proteomics_data %>% 
                  select(-orientation, -uniprotid, -locus)))
colnames(proteomics_t) <- proteomics_t[1, ]
proteomics_t <- proteomics_t[-1, ]
proteomics_t[, ] <- lapply(proteomics_t[,], as.numeric)
proteomics_t <- tibble::rownames_to_column(proteomics_t, var = "sample_id")
proteomics_t
```

```{r}
#| label: load-metabolomics
#| echo: false
#| include: true

# Not really sure how to use this, so I am ignoring this for now
metabolomics_data <- read.csv('metabolomics_data_cleaned.csv')

metabolomics_data 
```

```{r}
#| label: generate-model-data
#| echo: false
#| include: true
model_data <- clinical_edited %>%
  merge(proteomics_t, by = "sample_id")

model_data
```

```{r}
#| label: create-model-recipe
#| echo: false
#| include: false

# Recipe does all of the feature engineering so that we don't have to do anything else to the input data
model_recipe <- recipe(infection_status ~ ., data = model_data) %>%
  update_role(sample_id, new_role = "id") %>% # id role tells the model not to actually incorporate it as a feature
  update_role(patient_id, new_role = "id") %>%
  update_role(condition, new_role = "id") %>%
  update_role(infection_status, new_role = "outcome") %>% # specify target 
  step_dummy(all_nominal_predictors()) %>% # convert string to dummy variables
  step_normalize(all_numeric_predictors()) # normalize numeric features
```

```{r}
#| label: generate-train-test-split
#| echo: false
#| include: true
set.seed(42)
# Create train/test split
train_test_split <- initial_split(model_data,prop = 0.8, strata = infection_status)

train_data <- training(train_test_split)
test_data <- testing(train_test_split)
```

```{r}
#| label: create-model-object
#| echo: false
#| include: true

# Only doing a simple logistic regression since a tree-based model would most certainly overfit
log_reg <- logistic_reg(
  mixture = 1 # Lasso regression, using this to hopefully get rid of a lot of features that don't matter (sets coefficients of nonimportant features to 0, effectively removing them)
  penalty = tune() # testing a few different penalty values
) %>% 
  set_engine("glmnet") 
```

```{r}
#| label: generate-model-workflow
#| echo: false
#| include: true

# Create model workflow that includes model recipe and model object
model_workflow <- workflow() %>% 
  add_recipe(model_recipe) %>% 
  add_model(log_reg)
```

```{r}
#| label: generate-model-tuning-parameters
#| echo: false
#| include: true
set.seed(42)

# 5 fold cross validation
cv_folds <- vfold_cv(train_data, v = 5)

# Tune the penalty parameter
tune_res <- tune_grid(
  model_workflow,
  resamples = cv_folds,
  grid = 5,     # Try 5 values of the penalty
  metrics = metric_set(roc_auc, accuracy), # generate performance metrics for roc_auc & accuracy
  control = control_grid(save_pred = TRUE)
)
```

```{r}
#| label: view-model-metrics
#| echo: false
#| include: true

# ...almost all of them are really good?
tune_res %>% 
  collect_metrics()
```

```{r}
#| label: create-final-workflow
#| echo: false
#| include: true

# Show performance metrics and get the best model
best_result <- tune_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  slice_max(mean, n = 1) %>%
  head(1) # I have never had a tie for the best performing model, so I'm just grabbing the first one. Probably not the best strategy but this is new territory

# Finalize the workflow using the model workflow and our best performing model
final_workflow <- finalize_workflow(model_workflow, best_result)

# Fit the Final Model on the Training Data
fit_workflow <- fit(final_workflow, train_data)
```

```{r}
#| label: view-model-coefficients
#| echo: false
#| include: true

# Coefficients for features that seem to matter to the model, the rest have been set to 0 via the lasso regression
model_summary <- tidy(fit_workflow)
model_summary %>%
  filter(estimate != 0)
```

```{r}
#| label: test-set-predictions
#| echo: false
#| include: true

# Get predictions on test data
predictions <- predict(fit_workflow, test_data)
```

```{r}
#| label: bind-predictions-and-test-data
#| echo: false
#| include: true

# look at predictions on test data
results <- bind_cols(predictions, test_data)

results
```

```{r}
#| label: test-confusion-matrix
#| echo: false
#| include: true

# Generate confusion matrix on test data
confusion_matrix <- confusionMatrix(as.factor(results$.pred_class), as.factor(results$infection_status))

```

```{r}
#| label: view-confusion-matrix
#| echo: false
#| include: true

# The model is performing very well, just one misclassification
confusion_matrix$table
```