### 2024-11-05 04_shiny_app/Biomarker_Explorer/www/preparing_files.Rmd
Msystems resubmission

---
title: "fixing files"
output: html_document
date: "2024-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggpubr)
library(tidyverse)
```

```{r}
## Fixing the stats files ##
 protein_to_gene = readr::read_csv("../../../00_r_input_data/protein_id_to_gene.csv")
  gene_description = read_csv("../../../00_r_input_data/all_protein_annotations.csv")
  ## Loading all the data that will be needed ### 
    healthy = read_csv("../../../01_r_processed_data/Infected_vs_Healthy.csv") %>% 
      left_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
      left_join(gene_description, by = c("uniprotid" = "Entry")) %>% 
      mutate(url = paste0("https://www.uniprot.org/uniprotkb/",uniprotid,"/entry")) %>% 
      rename(Feature = Gene,
             Description = Protein.names)

    EcB = read_csv("../../../01_r_processed_data/faecalis_v_faecium.csv") %>% 
      left_join(protein_to_gene,by = c("uniprotid" = "ProteinID"))%>% 
      left_join(gene_description, by = c("uniprotid" = "Entry")) %>% 
      mutate(url = paste0("https://www.uniprot.org/uniprotkb/",uniprotid,"/entry")) %>% 
      rename(Feature = Gene,
             Description = Protein.names)

    mortality = read_csv("../../../01_r_processed_data/death_vs_life.csv") %>% 
      select(-Gene) %>% 
      left_join(protein_to_gene,by = c("uniprotid" = "ProteinID"))%>% 
      left_join(gene_description, by = c("uniprotid" = "Entry")) %>% 
      mutate(url = paste0("https://www.uniprot.org/uniprotkb/",uniprotid,"/entry")) %>% 
      rename(Feature = Gene,
             Description = Protein.names)

write_csv(healthy,"Infected_vs_Healthy.csv")
write_csv(EcB,"faecalis_v_faecium.csv")
write_csv(mortality,"death_vs_life.csv")

```

```{r}
## Fixing the continuous data files ##

data = read_csv("../../../01_r_processed_data/quant_msstats_protein_level_data.csv")

#Loading in mapping file
mapping = readxl::read_excel("../../../00_r_input_data/proteomics_sample_mapping.xlsx") %>% 
  mutate(sample_id = paste0("plex",plex_number,".",TMT_channel)) %>% 
  select(sample_id,sample_name)

# Loading in the clinical metadata 
cmd = read_csv("../../../00_r_input_data/clinical_metadata.csv")

mapping = inner_join(mapping,cmd,by = c("sample_name" = "sample_id"))

#Loading in data

data_md = inner_join(mapping,data,by = "sample_id")

#Converting into a matrix
f = data_md %>% 
  dplyr::select(sample_name,condition,death_during_admission,Gene,Abundance) %>% 
  dplyr::rename(Feature = Gene)

write_csv(f,"protein_data.csv")

```

# Generating stats for all the data

```{r}
gnps_annotations = read_csv("../../../00_r_input_data/gnps_annotations.csv")
normalized_data = read_csv("../../../01_r_processed_data/normalized_metabolomics.csv")

### Infected vs Healthy ###

  stat = normalized_data %>%
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ infection_status) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

  # Extracting the values for each feature for healthy patients
healthy = normalized_data %>% 
  filter(infection_status == "uninfected") %>% 
  group_by(row_id) %>% 
  summarise(healthy = mean(norm_value))

# Extracting the values for each feature for infected patients
infected = normalized_data %>% 
  filter(infection_status == "infected") %>% 
  group_by(row_id) %>% 
  summarise(infected = mean(norm_value))

## Combining the healthy and infected so we can caclulated the log2fc ##
log2fc = inner_join(healthy,infected,by = "row_id") %>% 
  mutate(log2fc_infected_healthy = log2(infected/healthy))

## Joining the statistics and log2 fold change ##  
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  left_join(gnps_annotations, by = "row_id") %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE)) %>% 
  select(row_id = row_id,Feature = compound_name_cleaned,log2FC = log2fc_infected_healthy,pvalue = p,adj.pvalue = p.adj_fdr,annotated) %>% 
  mutate(url = paste0("https://pubchem.ncbi.nlm.nih.gov/compound/",Feature)) %>% 
  mutate(url = case_when(url == "https://pubchem.ncbi.nlm.nih.gov/compound/NA" ~
                           "https://pubchem.ncbi.nlm.nih.gov/compound/no_name_in_data",
                         TRUE ~ url))

write_csv(vp,"healthy_vs_infected_metabolomics.csv")

### Faecalis vs Faecium ###
  stat = normalized_data %>%
    filter(condition %in% c("faecalis","faecium")) %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ condition) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

  # Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,condition,norm_value)

  # Extracting just the faecalis data
faecalis = log2fc_data %>% 
  filter(condition == "faecalis") %>% 
  group_by(row_id) %>% 
  summarise(faecalis = mean(norm_value))

# Extracting just the faecium data 
faecium = log2fc_data %>% 
  filter(condition == "faecium") %>% 
  group_by(row_id) %>% 
  summarise(faecium = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(faecalis,faecium,by = "row_id") %>% 
  mutate(log2fc_faecalis_faecium = log2(faecalis/faecium))

# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  left_join(gnps_annotations, by = "row_id") %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE)) %>% 
  select(row_id = row_id,Feature = compound_name_cleaned,log2FC = log2fc_faecalis_faecium,pvalue = p,adj.pvalue = p.adj_fdr,annotated) %>% 
  mutate(url = paste0("https://pubchem.ncbi.nlm.nih.gov/compound/",Feature)) %>% 
  mutate(url = case_when(url == "https://pubchem.ncbi.nlm.nih.gov/compound/NA" ~
                           "https://pubchem.ncbi.nlm.nih.gov/compound/no_name_in_data",
                         TRUE ~ url))

write_csv(vp,"faecalis_v_faecium_metabolomics.csv")

### Mortality vs Survival ###

    stat = normalized_data %>% 
    dplyr::group_by(row_id) %>%
    rstatix::t_test(norm_value ~ death_during_admission) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

    # Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,death_during_admission,norm_value)

# Extracting just the faecalis data
dead = log2fc_data %>% 
  filter(death_during_admission == TRUE) %>% 
  group_by(row_id) %>% 
  summarise(dead = mean(norm_value))

# Extracting just the faecium data 
live = log2fc_data %>% 
  filter(death_during_admission == FALSE) %>% 
  group_by(row_id) %>% 
  summarise(live = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(dead,live,by = "row_id") %>% 
  mutate(log2fc_live_dead = log2(live/dead))

# Combining log2fc and statistics 
# Combining log2fc and statistics 
vp = inner_join(stat,log2fc,by = "row_id")  %>% 
  left_join(gnps_annotations, by = "row_id") %>% 
  mutate(annotated = case_when(row_id %in% gnps_annotations$row_id ~ TRUE,
                               TRUE ~ FALSE)) %>% 
  select(row_id = row_id,Feature = compound_name_cleaned,log2FC = log2fc_live_dead,pvalue = p,adj.pvalue = p.adj_fdr,annotated) %>% 
  mutate(url = paste0("https://pubchem.ncbi.nlm.nih.gov/compound/",Feature)) %>% 
  mutate(url = case_when(url == "https://pubchem.ncbi.nlm.nih.gov/compound/NA" ~
                           "https://pubchem.ncbi.nlm.nih.gov/compound/no_name_in_data",
                         TRUE ~ url))

write_csv(vp,"death_vs_life_metabolomics.csv")

```

# Continuous data for metabolomics

```{r}

gnps_annotations = read_csv("../../../00_r_input_data/gnps_annotations.csv") %>% 
  select(row_id, compound_name_cleaned)

data = read_csv("../../../01_r_processed_data/normalized_metabolomics.csv") %>% 
  select(row_id,sample_id,condition,death_during_admission,Abundance = norm_value) %>% 
  inner_join(gnps_annotations, by = "row_id") %>% 
  rename(Feature = row_id) %>% 
  select(Feature,compound_name_cleaned, sample_name = sample_id, everything())

write_csv(data,"normalized_metabolomics.csv")
```