### 2024-11-05 08_Msystems_resubmission/01_Confounding_Analysis/resubmission_analysis.Rmd

---
title: "Resubmission Analysis"
author: "Charlie Bayne"
date: "`r Sys.Date()`"
output:
  html_document:
        theme: cosmo
        toc: true
        toc_depth: 6
        toc_float:
            collapsed: true
            smooth_scroll: true
        code_folding: hide
        highlight: tango
        df_print: paged
---

```{r setup, include=FALSE, warning= FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)

knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,cache = TRUE)

#Setting ggplot2 theme, this isn't working for some reason
theme_set(
  ggprism::theme_prism()+
  theme(plot.title = element_text(hjust = 0.5))
  )

#Setting ggplot2 theme, this one works
#theme_set(
 # theme_classic()+
  #theme(plot.title = element_text(hjust = 0.5))
  #)
```

# Reviewer 1 Comment 2

## Proteomics

```{r,fig.height = 4, fig.width = 7 }
infected_vs_healthy = readr::read_csv("../../01_r_processed_data/Infected_vs_Healthy.csv") %>% 
  dplyr::select(uniprotid,
                infected_v_healthy_log2fc = log2FC,
                infected_v_healthy_adj.p = adj.pvalue)

death_vs_life = readr::read_csv("../../01_r_processed_data/death_vs_life.csv") %>% 
  dplyr::select(uniprotid,
                death_vs_life_log2fc = log2FC,
                death_vs_life_adj.p = adj.pvalue) %>% 
  mutate(death_vs_life_log2fc = death_vs_life_log2fc * -1)

all = dplyr::inner_join(infected_vs_healthy,death_vs_life,by = "uniprotid") %>% 
  mutate(direction = case_when(death_vs_life_log2fc * infected_v_healthy_log2fc > 0 ~ "same_direction",
                                   TRUE ~ "different_direction"),
         sig = case_when((infected_v_healthy_adj.p <= 0.05 & death_vs_life_adj.p <= 0.05) ~ "both",
                         (infected_v_healthy_adj.p <= 0.05 & death_vs_life_adj.p > 0.05) ~ "infected_v_healthy",
                         (infected_v_healthy_adj.p > 0.05 & death_vs_life_adj.p <= 0.05) ~ "death_v_life",
                         TRUE ~ "neither")) %>% 
  filter(uniprotid != "Q6UY14")

lm = lm(data = all, infected_v_healthy_log2fc ~ death_vs_life_log2fc)
summary(lm)

p1 = all %>% 
  #filter(sig == "both") %>% 
  ggplot(aes(infected_v_healthy_log2fc,
             death_vs_life_log2fc))+
  geom_point(aes(color = direction))+
  geom_abline(slope = 1, linetype = "dashed")+
  geom_smooth(method = "lm")+
  ggtitle("All Proteins")+
  xlab("log2(Infected/Healthy)")+
  ylab("log2(Mortality/Survival)")

t1 = gridExtra::tableGrob(as.data.frame(table(all$direction)))

## Sig only ##

sig = all %>% 
  filter(sig == "both")

lm = lm(data = sig, infected_v_healthy_log2fc ~ death_vs_life_log2fc)
summary(lm)

p2 = sig %>% 
  #filter(sig == "both") %>% 
  ggplot(aes(infected_v_healthy_log2fc,
             death_vs_life_log2fc))+
  geom_point(aes(color = direction))+
  geom_smooth(method = "lm")+
  geom_abline(slope = 1, linetype = "dashed")+
  ggtitle("Proteins p.adj <= 0.05")+
  xlab("log2(Infected/Healthy)")+
  ylab("log2(Mortality/Survival)")
t2 = gridExtra::tableGrob(as.data.frame(table(sig$direction)))

p3 = ggpubr::ggarrange(p1,p2,common.legend = TRUE,legend = "bottom")

p3

ggsave("proteomics_correlation.pdf",p3, height = 4, width = 8)

ggsave("proteomics_table.pdf",t2)

```

## Metabolomics

```{r,fig.height =6, fig.width = 12}
## Healthy Metabolomics ## 
gnps_annotations = read_csv("../../00_r_input_data/gnps_annotations.csv",na = c("N/A","NA")) %>% 
  select(row_id,compound_name_cleaned)
## Reading in the normalized metabolomics data ##
normalized_data = read_csv("../../01_r_processed_data/normalized_metabolomics.csv")

normalized_data = inner_join(gnps_annotations,normalized_data,by = "row_id")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>%
    dplyr::group_by(row_id,compound_name_cleaned) %>%
    rstatix::t_test(norm_value ~ infection_status) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Extracting the values for each feature for healthy patients
healthy = normalized_data %>% 
  filter(infection_status == "uninfected") %>% 
  group_by(row_id,compound_name_cleaned) %>% 
  summarise(healthy = mean(norm_value))

# Extracting the values for each feature for infected patients
infected = normalized_data %>% 
  filter(infection_status == "infected") %>% 
  group_by(row_id,compound_name_cleaned) %>% 
  summarise(infected = mean(norm_value))

## Combining the healthy and infected so we can caclulated the log2fc ##
log2fc = inner_join(healthy,infected,by = "row_id") %>% 
  mutate(log2fc_infected_healthy = log2(infected/healthy))

## Joining the statistics and log2 fold change ##  
healthy = inner_join(stat,log2fc,by = "row_id")  %>%

  select(row_id,
         log2fc_infected_healthy,
         p.adj_fdr_health =p.adj_fdr)

### Survival vs Mortality ###

## Reading in Data
normalized_data = read_csv("../../01_r_processed_data/normalized_metabolomics.csv") %>% 
  filter(!is.na(death_during_admission))

normalized_data = inner_join(gnps_annotations,normalized_data,by = "row_id")

# Performing statistics so we can make a volcano plot
  stat = normalized_data %>% 
    dplyr::group_by(row_id,compound_name_cleaned) %>%
    rstatix::t_test(norm_value ~ death_during_admission) %>%
    rstatix::adjust_pvalue(p.col = "p",output.col = "p.adj_fdr", method = "fdr")

# Calculating log2fc
log2fc_data = normalized_data %>%  
    dplyr::select(metabolomics_id,row_id,compound_name_cleaned,death_during_admission,norm_value)

# Extracting just the faecalis data
dead = log2fc_data %>% 
  filter(death_during_admission == TRUE) %>% 
  group_by(row_id,compound_name_cleaned) %>% 
  summarise(dead = mean(norm_value))

# Extracting just the faecium data 
live = log2fc_data %>% 
  filter(death_during_admission == FALSE) %>% 
  group_by(row_id,compound_name_cleaned) %>% 
  summarise(live = mean(norm_value))

# Calculating log2fc
log2fc = inner_join(dead,live,by = "row_id") %>% 
  mutate(log2fc_dead_live = log2(dead/live))

# Combining log2fc and statistics 
mort = inner_join(stat,log2fc,by = "row_id")  %>% 
  select(row_id,
         log2fc_dead_live,
         p.adj_fdr_mort =p.adj_fdr,
         compound_name_cleaned)

all = inner_join(mort,healthy, by = "row_id") %>% 
    mutate(direction = case_when(log2fc_infected_healthy * log2fc_dead_live > 0 ~ "same_direction",
                                   TRUE ~ "different_direction")) %>% 
  mutate(sig = case_when((p.adj_fdr_mort <= 0.05 & p.adj_fdr_health <= 0.05) ~ "both",
                         (p.adj_fdr_health <= 0.05 & p.adj_fdr_mort > 0.05) ~ "infected_v_healthy",
                         (p.adj_fdr_health > 0.05 & p.adj_fdr_mort <= 0.05) ~ "death_v_life",
                         TRUE ~ "neither"))

expected_confounders = c("dibutyl phthalate","phthalic anhydride")

expected_true = c("cholic_acid",
                  "taurocholic acid",
                  "glycocholic acid",
                  "tauroursodeoxycholic acid",
                  "taurohyodeoxycholic acid")

labs = all %>% 
  filter(compound_name_cleaned %in% c(expected_true,expected_confounders))

lm = lm(data = all, log2fc_infected_healthy ~ log2fc_dead_live)
summary(lm)

p1 = all %>% 
  ggplot(aes(log2fc_infected_healthy,log2fc_dead_live))+
  geom_point(aes(color = direction))+
  geom_abline(slope = 1, linetype = "dashed")+
  ggrepel::geom_text_repel(data = labs, aes(label = compound_name_cleaned), size = 2)+
  geom_smooth(method = "lm")+
  ggtitle("All Annotated Metabolites") + 
  xlab("log2(Infected/Healthy)")+
  ylab("log2(Mortality/Survival)")

p1

t1 = gridExtra::tableGrob(as.data.frame(table(all$direction)))

sig = all %>% 
  filter(sig == "both")

lm = lm(data = sig, log2fc_infected_healthy ~ log2fc_dead_live)
summary(lm)

p2 = sig %>% 
  ggplot(aes(log2fc_infected_healthy,log2fc_dead_live))+
  geom_point(aes(color = direction))+
  geom_abline(slope = 1, linetype = "dashed")+
  ggrepel::geom_text_repel(aes(label = compound_name_cleaned), size = 2)+
  geom_smooth(method = "lm")+
  ggtitle("Metabolites p.adj <= 0.05")+
  xlab("log2(Infected/Healthy)")+
  ylab("log2(Mortality/Survival)")

p2

t2 = gridExtra::tableGrob(as.data.frame(table(sig$direction)))

p3 = ggpubr::ggarrange(p1,p2,common.legend = TRUE,legend = "bottom")

p3

ggsave("metabolomics_correlation.pdf",p3, height = 4, width = 8)

ggsave("metabolomics_table.pdf",t2)

```